---
subtitle: "0-Microarray data analysis"
date: "`r format (Sys.time(), format='%Y-%m-%d %H:%M:%S %z')`"
bibliography: bib.bib
output: 
  #bookdown::pdf_book:
  pdf_document:
    # base_format: rticles::springer_article
    pandoc_args:
      - '../common.yaml'
      - --listings
    includes: 
      in_header: '../preamble.tex'
    toc: true
    number_sections: true
    fig_width: 7
    fig_height: 6
    fig_caption: true
    df_print: tibble
---


```{r include=FALSE}
knitr::opts_chunk$set(tidy.opts=list(width.cutoff=80),
               tidy=TRUE,
               message = FALSE, 
               warning = FALSE)

```

\newpage

# Description

In order to build a lung interstitial macrophage (IM) specific mouse model, we need to find genes specific to this cell type comparing to other myeloid cell types in the lung. That's why we collected published microarray data and compared the expression profile of them. 

The following published microarray data were used. 

| Sample name                                               | Source         | IDENTIFIER (Raw data)                                                            |
|-----------------------------------------------------------|----------------|----------------------------------------------------------------------------------|
| Classical Monocytes MHCII+ in blood                       | ImmGen         | GEO: GSM605868, GSM605870, GSM605871                                             |
| Classical Monocytes MHCII- in bone marrow                 | ImmGen         | GEO: GSM854329, GSM854330, GSM854331                                             |
| Classical Monocytes MHCII- in blood                       | ImmGen         | GEO: GSM605872, GSM605873, GSM605874                                             |
| Nonclassical Monocytes, MHCII+                            | ImmGen         | GEO: GSM605878, GSM605879                                                        |
| Nonclassical Monocytes in bone marrow                     | ImmGen         | GEO: GSM854332, GSM854333, GSM854334                                             |
| Nonclassical Monocytes in blood                           | ImmGen         | GEO: GSM605884, GSM605885                                                        |
| Nonclassical Monocytes, MHCII int                         | ImmGen         | GEO: GSM605886, GSM605887, GSM605888, GSM605889, GSM605890                       |
| Lung CD11b+ CD24- macrophage                              | ImmGen         | GEO: GSM854271, GSM854272                                                        |
| Small Intestinal Lamina Propria CD11c-hi CD103- CD11b+ MF | ImmGen         | GEO: GSM854262, GSM854263, GSM854264, GSM854265, GSM854266, GSM854267, GSM854268 |
| Bone marrow macrophages                                   | ImmGen         | GEO: GSM854317, GSM854318, GSM854319                                             |
| Spleen Red Pulp macrophages                               | ImmGen         | GEO: GSM605853, GSM605854, GSM605855                                             |
| Peritoneal macrophage steady state                        | ImmGen         | GEO: GSM854294, GSM854295, GSM854296                                             |
| Peritoneal cavity macrophages steady state                | ImmGen         | GEO: GSM605850, GSM605851, GSM605852                                             |
| Medullary macrophages from skin draining lymph nodes      | ImmGen         | GEO: GSM854322, GSM854323                                                        |
| Central nervous system microglia                          | ImmGen         | GEO: GSM854326, GSM854327, GSM854328                                             |
| CD103+ migratory DC, Mediastinal LN CD103+ DC             | ImmGen         | GEO: GSM854243, GSM854244, GSM854245                                             |
| CD11b+ migratory DC, Mediastinal LN CD11b+ DC             | ImmGen         | GEO: GSM854255, GSM854256, GSM854257                                             |
| Lung CD103+ dendritic cells                               | ImmGen         | GEO: GSM538231, GSM538232, GSM538233, GSM854241, GSM854242                       |
| Lung MHCII+ CD11c+ CD103- CD11b+ CD24+ dendritic cells    | ImmGen         | GEO: GSM854269, GSM854270                                                        |
| Lung IM, Ly6C+ cMo and AM                                 | [@Sabatel2017] | EMBL-EBI: E-MTAB-5012                                                            |


# Load packages and data

The gene expression activity matrix for each samples were generated in Gene Expression Commons (https://gexc.riken.jp) [@Seita2012]. 

```{r}
# packages
library(ComplexHeatmap)
library(RColorBrewer)
library(circlize)

# data
dir.files <- "./downloaded_data"
files.list <- list.files(dir.files, pattern = "Gene_Expression_Activity.csv", full.names = T)
names.list <- sub(basename(files.list), pattern = "_in_Gene_Expression_Activity.csv", replacement = "")
```

read csv files and bind tables to one: 

```{r}
expr.table <- data.frame(read.csv(files.list[1]), row.names = 1)
expr.table <- expr.table[, -2]
n.rep <- length(2:ncol(expr.table))
  
names.rep <- paste(rep(names.list[1]), 1:n.rep , sep = "_")
colnames(expr.table)[2:ncol(expr.table)] <- names.rep

for (i in 2:length(files.list)) {
  tb <- read.csv(files.list[i])
  tb <- tb[, 4:ncol(tb)]
  n <- ncol(tb)
  repname <- paste(rep(names.list[i]), 1:n, sep = "_")
  colnames(tb) <- repname
  
  expr.table <- cbind(expr.table, tb)
  n.rep <- append(n.rep, n)
  names.rep <- append(names.rep, repname)
  }
meta.sample <- data.frame(sampleName=names.list, 
                          n.rep=n.rep)
head(expr.table)
```


# Data preparration

## Make a list with genes to show in heatmap

>The gene list IMvs(AM&DC).csv is calculated in ImmGen Datasets. We compared IM microarray data to both MA and DC and get the DE genes. 

```{r}
probset.DE <- read.csv("./IMvs(AM&DC).csv")
probset.toShow <- unique(as.character(probset.DE$ProbeSet_ID))

# the table is ordered by ratio, so take the top 100: 
probset.toShow <- probset.toShow[1:100]
```

>Then base on the intensity in IM, we choose the only top 50 probsets. 

Take the to 50 probsets with highest intensity:

```{r}
probset.DE <- probset.DE[order(probset.DE$Mean_A, decreasing = TRUE), ]
probset.top50 <- as.character(probset.DE[1:50, "ProbeSet_ID"])
```

```{r}
probset.toShow <- intersect(probset.top50, probset.toShow)
```


subset expr.table

```{r}
expr.table.toShow <- expr.table[ probset.toShow, ]
genes.toShow <- unique(expr.table.toShow$Gene.Symbol)
length(genes.toShow)
```
*As genes are unique to each probset, we can use gene symbols as rownames. *

```{r}
rownames(expr.table.toShow) <- expr.table.toShow$Gene.Symbol
rownames(expr.table.toShow)
```

Remove the row without annotation:

```{r}
expr.table.toShow <- expr.table.toShow[expr.table.toShow$Gene.Symbol != "---", ]
```


## Make metadata table

```{r}
data.frame(meta.sample$sampleName, order=1:nrow(meta.sample))
```

```{r}
meta.sample$cellType3 <- c("Mac Alv Lu", #1
                           "DC CD103+ Lu", #2
                           "DC CD24+ Lu", #3
                           "DC CD103+ LuLN", #4
                           "DC CD11b+ LuLN", #5
                           "Mo Ly6C+ Lu", #6
                           "Mac Int Lu", #7
                           "Mac BM", #8
                           "Mac CNS", #9
                           "Mac Int Lu", #10
                           "Mac F4/80hi PC", #11
                           "Mac F4/80lo PC", #12
                           "Mac SI", #13
                           "Mac SLN", #14
                           "Mac SP", #15
                           "Mo Ly6C- MHCII- BL", #16
                           "Mo Ly6C- MHCII+ BL", #17
                           "Mo Ly6C- MHCIIint BL", #18
                           "Mo Ly6C+ MHCII- BL", #19
                           "Mo Ly6C+ MHCII+ BL", #20
                           "Mo Ly6C- MHCII- BM", #21
                           "Mo Ly6C+ MHCII- BM" #22
              )

meta.sample$cellType <- c("aMac", #1
              rep("DC", 4),  # 2-5
              "Mo", # 6
              "iMac", #7
              rep("Mac", 2), # 8-9 
              "iMac", #10
              rep("Mac", 5), # 11-15 
              rep("Mo", 7) # 16-22
              )
meta.sample$organ <- c(rep("Lu", 3), #1-3
           rep("LuLN", 2), #4-5
           rep("Lu", 2), #6-7
           "BM", #8
           "CNS", #9
           "Lu", #10
           rep("PC",2), #11-12
           "SI", #13
           "SLN", #14
           "SP", #15
           rep("BL",5 ), 
           rep("BM", 2)
           
           )
meta.sample$cellType2 <- c(
              "Mac", #1
              rep("DC", 4),  # 2-5
              "Mo", # 6
              "Mac", #7
              rep("Mac", 2), # 8-9 
              "Mac", #10
              rep("Mac", 5), # 11-15 
              rep("Mo", 7) # 16-22
              )
meta.sample$organ2 <- c(
            "Lu-Alv", #1
            rep("Lu", 2), #2-3
           rep("LuLN", 2), #4-5
           "Lu", #6
           "Lu-Int", #7
           "BM", #8
           "CNS", #9
           "Lu-Int", #10
           rep("PC",2), #11-12
           "SI", #13
           "SLN", #14
           "SP", #15
           rep("BL",5 ), 
           rep("BM", 2)
           
           )
```

```{r}
meta.table <- data.frame(CellType=rep(meta.sample$cellType, meta.sample$n.rep), 
                         OrganType=rep(meta.sample$organ, meta.sample$n.rep), 
                         CellType2=rep(meta.sample$cellType2, meta.sample$n.rep), 
                         OrganType2=rep(meta.sample$organ2, meta.sample$n.rep), 
                         cellType3=rep(meta.sample$cellType3, meta.sample$n.rep), 
                         row.names = names.rep)
```


```{r}
HeatmapAnnotation(Cell_type=meta.table$CellType, 
                  Organ_type=meta.table$OrganType)
```

# Make heatmaps

Use Heatmap: 

```{r fig.width=9, fig.height=7}
Heatmap(
  as.matrix(expr.table.toShow[2:ncol(expr.table.toShow)]),
  # use_raster = FALSE, # use FALSE to export to vector image. 
  name                         = "z-score",
  # col                          = colorRamp2(seq(from=-2,to=2,length=11),rev(brewer.pal(11, "Spectral"))),
  # show_row_names               = TRUE,
  # show_column_names            = FALSE,
  row_names_gp                 = gpar(fontsize = 7), 
  
  # row_title_rot                = 0,
  # cluster_rows                 = TRUE,
  # cluster_row_slices           = FALSE,
  #cluster_columns              = FALSE 
)
```

Using pheatmap. 

```{r fig.width=9, fig.height=7}
pheatmap(
  as.matrix(expr.table.toShow[2:ncol(expr.table.toShow)]), annotation_col = meta.table
  # use_raster = FALSE, # use FALSE to export to vector image. 
  #name                         = "z-score",
  # col                          = colorRamp2(seq(from=-2,to=2,length=11),rev(brewer.pal(11, "Spectral"))),
  # show_row_names               = TRUE,
  # show_column_names            = FALSE,
  #row_names_gp                 = gpar(fontsize = 7), 
  
  # row_title_rot                = 0,
  # cluster_rows                 = TRUE,
  # cluster_row_slices           = FALSE,
  #cluster_columns              = FALSE 
)
```

Using Heatmap with annotations. 
```{r fig.width=9, fig.height=7}
hp <- Heatmap(
  as.matrix(expr.table.toShow[2:ncol(expr.table.toShow)]),
  # use_raster = FALSE, # use FALSE to export to vector image. 
  name                         = "z-score",
  # col                          = colorRamp2(seq(from=-2,to=2,length=11),rev(brewer.pal(11, "Spectral"))),
  # show_row_names               = TRUE,
  # show_column_names            = FALSE,
  row_names_gp                 = gpar(fontsize = 7), 
  column_names_gp              = gpar(fontsize = 7), 

  #column_split = meta.table$CellType2, 
  column_split = factor(meta.table$CellType2, levels = c("Mac", "Mo", "DC")),
  top_annotation = HeatmapAnnotation(Organtype=meta.table$OrganType2, 
                                     col = list(Organtype = c(`Lu-Alv`="#32a852", 
                                             `Lu-Int`="#87c22f", 
                                             Lu="#205c30", 
                                             LuLN="#265d69", 
                                             BM="#82622f", 
                                             CNS="#4674e8", 
                                             PC="#a14bab", 
                                             SI="#dbed4e", 
                                             SLN="#ffa6f9", 
                                             SP="#2000f2", 
                                             BL="#f20000")) )
  
)

p <- draw(hp)
```

Change colors and annotations

```{r fig.width=9, fig.height=7}

col.cellType3 <- read.csv("../0-Microarrays/colors_celltype3.csv", sep = "\t", header = FALSE, row.names = 1)
colors.cellType3 <- as.character(col.cellType3$V2)
names(colors.cellType3) <- rownames(col.cellType3)
genes.toSplit <- rownames(expr.table.toShow)
genes.toSplit <- genes.toSplit %in% c("Tmem119", "Cx3cr1")

# the one with row split: 
hp2 <- Heatmap(
  as.matrix(expr.table.toShow[2:ncol(expr.table.toShow)]),
  # use_raster = FALSE, # use FALSE to export to vector image. 
  name                         = "Gene activity",
  row_names_gp                 = gpar(fontsize = 7), 
  column_names_gp              = gpar(fontsize = 7), 
  column_split = factor(meta.table$CellType2, levels = c("Mac", "Mo", "DC")),
  row_split = 2, 
  top_annotation = HeatmapAnnotation(Cell_type=meta.table$cellType3,
                    col = list(
                                Cell_type = colors.cellType3 ), 
                    annotation_legend_param = list(
                                Cell_type = list(title = "Cell type", 
                                                 at = names(colors.cellType3), 
                                                 labels = names(colors.cellType3))) )
  
)

p2 <- draw(hp2)
```

```{r fig.width=9, fig.height=7}
# the one WITHOUT row split: 
hp3 <- Heatmap(
  as.matrix(expr.table.toShow[2:ncol(expr.table.toShow)]),
  # use_raster = FALSE, # use FALSE to export to vector image. 
  name                         = "Gene activity",
  row_names_gp                 = gpar(fontsize = 7), 
  column_names_gp              = gpar(fontsize = 7), 
  column_split = factor(meta.table$CellType2, levels = c("Mac", "Mo", "DC")),
  #row_split = 2, 
  top_annotation = HeatmapAnnotation(Cell_type=meta.table$cellType3,
                    col = list(
                                Cell_type = colors.cellType3 ), 
                    annotation_legend_param = list(
                                Cell_type = list(title = "Cell type", 
                                                 at = names(colors.cellType3), 
                                                 labels = names(colors.cellType3))) )
  
)

p3 <- draw(hp3)
```


```{r include=FALSE}
# export data

pdf(file = "./plots/heatMap_top50_probsets.pdf", width = 9, height = 7)
print(p)
dev.off()

write.csv(expr.table.toShow, file = "./top50_exprTable.csv")

# export data

pdf(file = "./plots/heatMap_top50_probsets_with_21celltypes_with_rowSplit.pdf", width = 9, height = 7)
print(p2)
dev.off()

# export data

pdf(file = "./plots/heatMap_top50_probsets_with_21celltypes_without_rowSplit.pdf", width = 9, height = 7)
print(p3)
dev.off()
```

# Session information 

R sesssion: 

```{r}
sessionInfo()
```

# References

