---
subtitle: "13 - Compare with Mafb-KO IM"
# bibliography: bib.bib
link-citations: yes
date: "`r format (Sys.time(), format='%Y-%m-%d %H:%M:%S %z')`"
output: 
  #html_document
  pdf_document:
    pandoc_args:
      - '../common.yaml'
      - --listings
    includes:
      in_header: '../preamble.tex'
    toc: true
    number_sections: true
    fig_width: 7
    fig_height: 6
    fig_caption: true
    df_print: tibble
---

```{r include=FALSE}
knitr::opts_chunk$set(message = FALSE, warning = FALSE)
```

\newpage

# Description

In this analysis, we compared Mafb-KO sample to Control sample. The analysis peppine was the same as previous ones. We found new Mafb-KO specific population (population "Enriched in Mafb-KO"). Using signature scoring, we showed the "Enriched in Mafb-KO" population had low IM signature score than Control IM population but had higher cMo signature, indicating they were in a immature differeniation state. 

The IM and cMo signature scoring was made with VISION package [@DeTomaso2019]. Briefly, The IM- and cMo-specific gene signatures were calculated with previously published scRNAseq data [@Schyns2019] by comparing either IM or cMo population to all other cell types in the dataset using FindMarker function (Seurat). The genes with logFC > 1 and only positively regulated ones were considered as IM or cMo signature. The signatures were then used to calculate the scores for each cells in the IM and Mafb KO-enriched clusters with VISION package. The scores were stored in Seurat object and plotted with Seurat package.

GSEA results were obtained using GSEA desktop [@Subramanian2005] with the input data described in the following codes. The results could be found in the same analysis folder. 

# Load packages and data

```{r}
suppressMessages({
  library(Seurat)
  library(ggplot2)
  library(multcomp)
})

results <- readRDS(file = "../12-cMAF_and_Mafb_deficient_IM/All_samples_Maf.seuratObject.Rds")

# we will work on control and Mafb-KO samples, so subset: 
so <- subset(results, subset = group == c("HT5-Control", "HT7-MAFb-KO"))
```


# Re process data with only Control and Mafb-KO samples

## Re-process data

```{r}
so <- NormalizeData(so, verbose=FALSE)
so <- FindVariableFeatures(so, selection.method = "vst", nfeatures = 1000, verbose=FALSE) # we focus less variable genes. 
so <- ScaleData(so, features = rownames(so), verbose=FALSE)
so <- RunPCA(so, features = VariableFeatures(so), verbose=FALSE)
so <- RunTSNE(so, dims = 1:8, verbose=FALSE)
so <- RunUMAP(so, dims = 1:8, verbose=FALSE)
```

```{r message=FALSE, warning=FALSE}
so <- FindNeighbors(so, dims = 1:8, verbose = FALSE)
so <- FindClusters(so, resolution = 0.15, verbose = FALSE) 
```

```{r}
DimPlot(so, split.by = "group")
```


# Compare population composition

```{r, fig.height=4, fig.width=6}
pal3 <- c(
  "#A6CEE3", # cMo
  "#1F78B4", # pMo
  "#FF7F00", # Intermediate
  "#B2DF8A", # MHCII IM
  "#33A02C", # CD206 IM
  "#CAB2D6", # Mafb- neo
  "#ededed"# Unknown
)
DimPlot(so, cols = pal3[1:4])
```

```{r fig.height=16, fig.width=16}
VlnPlot(
  so, 
  features = c("Ccr2", "Plac8" ,"Ly6c2", "Spn", 
                              "Fcgr4", "Ace", "Cxcr4", 
                              "Fcgr1", "Mrc1", "Cd63", "Apoe", 
                              "H2-DMb2", "Cd74", "Cd209a", 
                              "Ccnd1","Mki67", "Top2a", "Ccl2", 
                              "Irf7", "Cxcl10", "Ifit1", "percent.mt", "Ifi47", "Itgax") , 
        ncol = 4, cols = pal3[1:4] )
```

cluster 0: Mafb-independent
Cluster 1: IM
Cluster 2: Classical Monocytes
Cluster 3: Patrolling Monocytes

```{r}
so$cell.type2 <- factor(Idents(so), labels = c("Mafb-independent", "IM", "Classical Mono", "Patrolling Mono"))
so$cell.type2 <- factor(so$cell.type2, levels = c("Classical Mono", "Patrolling Mono", "IM", "Mafb-independent"))
Idents(so) <- "cell.type2"
```

```{r}
pal4 <- c(
  "#A6CEE3", # cMo
  "#1F78B4", # pMo
  "#33A02C", # CD206 IM
   "#FF7F00" # Mafb- neo
)
DimPlot(so, cols = pal4, split.by = "group"
)
```
```{r fig.height=5, fig.width=7}
DimPlot(so, label = T, cols = pal4)
```
```{r eval=FALSE}
ggsave(filename = "../Figures/UMAPplot_Ctl_AND_MafbKO_with_legend.pdf", 
       width = 7, height =5)
```

Plot cell in colors but only for one of two samples

```{r}
so$cell.type.control <- as.character(so$cell.type2)
so$cell.type.control[WhichCells(so, expression = group == "HT7-MAFb-KO")] <- "Mafb-KO"
so$cell.type.control <- factor(so$cell.type.control, levels = c("Classical Mono", "Patrolling Mono", "IM", "Mafb-independent", "Mafb-KO"))
```

```{r fig.height=5, fig.width=7}
pal4.control <- c(
  "#F1F1F1", # grey for another sample
  "#FF7F00", # Mafb- neo
  "#A6CEE3", # cMo
  "#1F78B4", # pMo
  "#33A02C" # CD206 IM
)
DimPlot(so, cols = pal4.control, group.by = "cell.type.control", pt.size = 2,
order = c("IM", "Patrolling Mono", "Classical Mono", "Mafb-independent",  "Mafb-KO")
)
```


```{r eval=FALSE}
ggsave(filename = "../Figures/UMAPplot_Ctl_IM_with_legend.pdf", 
       width = 7, height =5)
```

```{r fig.height=5, fig.width=5}
DimPlot(so, cols = pal4.control, group.by = "cell.type.control", pt.size = 2,
order = c("IM", "Patrolling Mono", "Classical Mono", "Mafb-independent",  "Mafb-KO")
) + NoLegend()
```

```{r eval=FALSE}
ggsave(filename = "../Figures/UMAPplot_Ctl_(vsMafbKO)_no_legend.pdf", 
       width = 5, height = 5)
```

Plot Mafb-KO: 

```{r}
so$cell.type.mafbko <- as.character(so$cell.type2)
so$cell.type.mafbko[WhichCells(so, expression = group == "HT5-Control")] <- "Control"
so$cell.type.mafbko <- factor(so$cell.type.mafbko, levels = c("Classical Mono", "Patrolling Mono", "IM", "Mafb-independent", "Control"))
```

```{r fig.height=5, fig.width=5}
pal4.control <- c(
  "#F1F1F1", # grey for another sample
  "#FF7F00", # Mafb- neo
  "#A6CEE3", # cMo
  "#1F78B4", # pMo
  "#33A02C" # CD206 IM
)
DimPlot(so, cols = pal4.control, group.by = "cell.type.mafbko", pt.size = 2,
order = c("IM", "Patrolling Mono", "Classical Mono", "Mafb-independent",  "Mafb-KO")
) + NoLegend()
```

```{r eval=FALSE}
ggsave(filename = "../Figures/UMAPplot_MafbKO_(vsControl)_no_legend.pdf", 
       width = 5, height = 5)
```

See population frequencies:

```{r fig.width=3, fig.height=4}
source("../R/SeuratFreqTable.R")
freq.celltype.list <- list(
  `Control` = Seurat2CellFreqTable(subset(so, subset = group == "HT5-Control"), slotName = "cell.type2"),  
  `MAFb-KO` = Seurat2CellFreqTable(subset(so, subset = group == "HT7-MAFb-KO"), slotName = "cell.type2") 
)

source("../R/barChart.R")
barChart(freq.celltype.list) + labs(fill = "Cluster") + scale_fill_manual(values = pal4) + theme(axis.text.x = element_text(angle = 90))
```

```{r eval=FALSE}
ggsave(filename = "../Figures/Barplot_Ctl_MafbKO_population_frequency.pdf", 
       width = 3, height = 4)
```

Show CD206+ IM markers


```{r fig.height=4, fig.width=5}
FeaturePlot(so, features = "Mrc1", cols = c("lightgray", "red"), min.cutoff = 0.5)
```
```{r eval=FALSE}
ggsave(filename = "../Figures/VlnPlot_Mrc1_in_all_samples.pdf", width = 5, height = 4)
```

```{r fig.height=4, fig.width=5}
FeaturePlot(so, features = "Lyve1", cols = c("lightgray", "red"))
```
```{r eval=FALSE}
ggsave(filename = "../Figures/VlnPlot_Lyve1_in_all_samples.pdf", width = 5, height = 4)
```

```{r fig.height=8, fig.width=8}
FeaturePlot(so, features = c("Mrc1", "Lyve1"), cols = c("lightgray", "red"), split.by = "group")
```


# Proliferation comparison

```{r fig.width=3, fig.height=4}
freq.celltype.list <- list(
  `Control` = Seurat2CellFreqTable(subset(so, subset = group == "HT5-Control"), slotName = "Phase"),  
  `MAFb-KO` = Seurat2CellFreqTable(subset(so, subset = group == "HT7-MAFb-KO"), slotName = "Phase") 
)

barChart(freq.celltype.list) + labs(fill = "Cluster") +  theme(axis.text.x = element_text(angle = 90))
```

```{r fig.width=9, fig.height=4}
DimPlot(so, group.by = "Phase", split.by = "group")
```

```{r}
VlnPlot(so, features = "G2M.Score", split.by = "group", cols = pal3, split.plot = TRUE)
```

# Different population between Mafb-KO and Control sample

Let's focus on the Mafb-deficient population in Mafb-deficient sample.

```{r}
neo_IM <- subset(so, subset = cell.type2 %in% c("Mafb-independent", "IM"))
```

## DE genes between Mafb- neo and IM population

```{r fig.width=8, fig.height=10}
library(dplyr)
Neo_vs_IM <- FindMarkers(so, 
                          ident.1 = "Mafb-independent", 
                          ident.2 = c("IM"), 
                          logfc.threshold = 0,
                          verbose = FALSE)


# keep only adj p value < 0.05 and logFC > 0.5 as significant markers. 
Neo_vs_IM.markers <- Neo_vs_IM[Neo_vs_IM$p_val_adj < 0.05 & abs(Neo_vs_IM$avg_log2FC) > 0.5, ]
Neo_vs_IM.markers <- Neo_vs_IM.markers[order(Neo_vs_IM.markers$avg_log2FC, decreasing = TRUE), ]
nrow(Neo_vs_IM.markers)
```

```{r eval=FALSE}
write.csv(Neo_vs_IM.markers ,file = "./Mafb-deficient_vs_IM.DEgenes.results.csv", quote = FALSE)
```


Let's show the top 20 of each side: 

```{r}
Neo_vs_IM.markers.top20 <- Neo_vs_IM.markers[c(1:20, (nrow(Neo_vs_IM.markers)-19):(nrow(Neo_vs_IM.markers))), ]
```


```{r message=FALSE}
library(ComplexHeatmap)
mat <- GetAssayData(neo_IM)[rownames(Neo_vs_IM.markers.top20), ]
mat.scale <- t(scale(t(as.matrix(mat))))
```

```{r fig.height=5, fig.width=6}
hp <- Heatmap(mat.scale, show_row_names = TRUE, show_column_names = FALSE, 
        row_names_gp = gpar ( fontsize = 7), 
        column_split = factor(neo_IM$cell.type2), 
        km = 2)
hp <- draw(hp)
```


```{r eval=FALSE}
pdf(file = "../Figures/Heatmap_IM_vs_Mafb-neo.pdf", width = 6, height = 5)
hp
dev.off()
```

Show in vlnplot

```{r fig.height=8, fig.width=8}
p <- VlnPlot(neo_IM, features = c("Mrc1", "Adgre1", "Pf4", "Tmem176a",  "Tmem176b","Tmem119", "Apoe", "C1qa", "C1qb", "C1qc", "Mafb", "Cd63"), group.by = "cell.type2", cols = c("#33A02C", "#FF7F00"), ncol = 4, pt.size = 0)
p
```

Show vlnplot without label: 

```{r fig.height=7, fig.width=8}
p & theme(axis.title.x=element_blank(),
        axis.text.x=element_blank(),
        axis.ticks.x=element_blank())
```
```{r eval=FALSE}
ggsave(filename = "../Figures/Vlnplot_IMID_genes_in_Ctrl_MafbKO.pdf", 
       width = 8, height = 8)
```


## GO enrichment analysis with DE genes

```{r}
suppressMessages(library(clusterProfiler))
source("../R/entrez2symbol.R")
source("../R/replaceEntrezID.R")
```

### GO enrichment analysis of up-regulated DE genes in Mafb-deficient population

```{r message=FALSE}
DE.MafbKO <- Neo_vs_IM.markers[Neo_vs_IM.markers$avg_log2FC > 0, ]
symb <- rownames(DE.MafbKO)
de_entrez <- bitr( geneID = symb, fromType = "SYMBOL", toType = "ENTREZID", OrgDb = "org.Mm.eg.db", drop = TRUE ) $ ENTREZID
result.enrichGO <- enrichGO(de_entrez, OrgDb = "org.Mm.eg.db", ont = "BP")
result.enrichGO <- replaceEntrezID(result.enrichGO, organism = "mmu")
# write.csv(result.enrichGO@result, file = "./Results_enrichment/enrichGO_DE_Mafb_KO_vsIM.csv")
result.enrichGO@result
```

### GO enrichment analysis of up-regulated DE genes in IMs

```{r message=FALSE}
DE.IM <- Neo_vs_IM.markers[Neo_vs_IM.markers$avg_log2FC < 0, ]
symb <- rownames(DE.IM)
de_entrez <- bitr( geneID = symb, fromType = "SYMBOL", toType = "ENTREZID", OrgDb = "org.Mm.eg.db", drop = TRUE ) $ ENTREZID
result.enrichGO <- enrichGO(de_entrez, OrgDb = "org.Mm.eg.db", ont = "BP")
result.enrichGO <- replaceEntrezID(result.enrichGO, organism = "mmu")
# write.csv(result.enrichGO@result, file = "./Results_enrichment/enrichGO_DE_IM_vsMafbKO.csv")
result.enrichGO@result
```

### Volcano plot of DE genes

```{r}
suppressMessages({
  library(dplyr)
  library(ggrepel)
  })
```


Let's set a threshold of log2FC and p_val_adj and plot them all: 

```{r fig.height=5, fig.width=6}
threshold.log2fc <- 0.5
threshold.adjp <- 0.05
Neo_vs_IM.volcano = mutate(Neo_vs_IM, 
 Sig=ifelse((abs(Neo_vs_IM$avg_log2FC) > threshold.log2fc)&(Neo_vs_IM$p_val_adj < threshold.adjp), "Sig", "n.s."))
# add two colors to 2 sig lists
Neo_vs_IM.volcano$Sig [ Neo_vs_IM.volcano$avg_log2FC < -threshold.log2fc & Neo_vs_IM.volcano$p_val_adj < threshold.adjp ] <- "1"

Neo_vs_IM.volcano$Sig [ Neo_vs_IM.volcano$avg_log2FC > threshold.log2fc & Neo_vs_IM.volcano$p_val_adj < threshold.adjp ] <- "2"

Neo_vs_IM.volcano$Gene <- rownames(Neo_vs_IM.volcano)
Gene.to.show.ValcanoPlot <- rownames(Neo_vs_IM.volcano[Neo_vs_IM.volcano$Sig != "n.s.", ])

p <- ggplot(Neo_vs_IM.volcano, aes(avg_log2FC, -log10(p_val_adj))) + geom_point(aes(col=Sig)) + scale_color_manual(values=c( `1`="#33A02C", `n.s.`="grey", `2`="#FF7F00")) 

# set axis lim: 
axis.lim <- max(abs(Neo_vs_IM.volcano$avg_log2FC)) + 1

p + geom_text_repel(data=filter(Neo_vs_IM.volcano, Gene %in% Gene.to.show.ValcanoPlot), size = 8, aes(label=Gene, fontface = "italic"), box.padding = 1) + xlim(c(-axis.lim, axis.lim)) + theme_classic() + theme(panel.border = element_rect(colour = "black", fill = NA, size = 1)) + geom_hline(yintercept = -log10(threshold.adjp), linetype='dashed', col = 'grey') + geom_vline(xintercept = c(-threshold.log2fc, threshold.log2fc),  linetype='dashed', col = 'grey') + ggtitle(paste("Log2FC > ", threshold.log2fc, "; p_val_adj <", threshold.adjp))

# write.csv(Neo_vs_IM.volcano[Gene.to.show.ValcanoPlot, ] %>% arrange(., desc(avg_log2FC)) , 
#          file = paste("./Mafb-deficient_vs_IM.DEgenes.Log2FC", threshold.log2fc, ".adjPval", threshold.adjp, ".results.csv", sep = "") ) 

```
```{r eval=FALSE}
ggsave(filename = "../Figures/VolcanoPlot_DE_IM_ctrol_vs_MafbKO.pdf", 
       width = 6, height = 5)
```


# GSEA analysis: Mafb-KO population vs IM


```{r eval=FALSE}
expr.table <- GetAssayData(neo_IM, slot = "data", assay = "RNA")
expr.table <- as.data.frame(expr.table)
row.info <- data.frame(NAME=rownames(expr.table), DESCRIPTION=rownames(expr.table))
expr.table <- cbind(row.info, expr.table)

# write.table(expr.table, file = "./AssayData_RNA_data_MafbKO_control_IM.txt", sep = "\t", quote = FALSE, row.names = FALSE)

# metadata: 
cls.table <- matrix(as.character(neo_IM@meta.data$cell.type2), nrow = 1)
# write.table(cls.table, file = "./Class_celltype2_MafbKO_control_IM.cls", sep = " ", quote = FALSE, col.names = FALSE, row.names = FALSE) # pay attention to the row.names=FALSE, or it will add row 1 to the head. 
```

```{r eval=FALSE}
# To add to cls file: 
cat(paste(length(cls.table), length(unique(as.character(cls.table))), 1), 
    "\n", 
    "# ", paste(unique(as.character(cls.table)), collapse = " "), 
    sep = "")
```

```{r eval=FALSE}
# For chip file
all.features <- read.csv(file = "../../../IM-DTR_MAF/counts/scRNAseq/Experiment-7-12-21-ScRNA_NGS21-U976/outs/raw_feature_bc_matrix/features.tsv.gz", sep = "\t")[, 2]
chip.file <- data.frame(`Probe Set ID`=all.features, 
                        `Gene Symbol`=all.features, 
                        `Gene Title`=all.features)
# write.table(chip.file, file = "./genename.chip",sep = "\t", quote = FALSE, row.names = FALSE )
```

The results can be found in sub directory: GSEA

# Scoring of IM and monocyte signatures in control and Mafb-KO samples


## Load data

We start from our published data [@Schyns2019].

The object "IMss_monocytes_filtered.v3" is the Seurat object used in this report. 

```{r include=FALSE}
load("/media/qiang/Data2/Archive/0_Archived_Projects/2019_Joey/20190509_Seurat_objects_for_SCENIC/IMss_monocytes_filtered.v3")
```

```{r}
suppressMessages(library(VISION))
Idents(IMss_monocytes_filtered.v3) <- "population"
DimPlot(IMss_monocytes_filtered.v3) + ggtitle("scRNAseq data in Schyns et al. 2019 Nat Comm")
```

## Create signautres for IM, classical monocytes and patrolling monocytes

```{r}
signature.im <- FindMarkers(IMss_monocytes_filtered.v3, 
                            ident.1 = c("CD206 IM", "MHC-II IM"), 
                            only.pos = TRUE, 
                            logfc.threshold = 1, verbose = FALSE)
signature.pmo <- FindMarkers(IMss_monocytes_filtered.v3, 
                            ident.1 = "Patrolling Mono", 
                            only.pos = TRUE, 
                            logfc.threshold = 1, verbose = FALSE)
signature.cmo <- FindMarkers(IMss_monocytes_filtered.v3, 
                            ident.1 = "Classical Mono", 
                            only.pos = TRUE, 
                            logfc.threshold = 1, verbose = FALSE)
```

```{r}
sig.im <- rep(1, length(rownames(signature.im)))
names(sig.im) <- rownames(signature.im)
sig.im <- createGeneSignature(name = "IM", sigData = sig.im)

sig.cmo <- rep(1, length(rownames(signature.cmo)))
names(sig.cmo) <- rownames(signature.cmo)
sig.cmo <- createGeneSignature(name = "cMo", sigData = sig.cmo)

sig.pmo <- rep(1, length(rownames(signature.pmo)))
names(sig.pmo) <- rownames(signature.pmo)
sig.pmo <- createGeneSignature(name = "pMo", sigData = sig.pmo)

sig.im_mono <- c(sig.im, sig.cmo, sig.pmo)
```

## Signature scoring

```{r message=FALSE}
vis <- Vision(so, signatures = sig.im_mono)

vis <- calcSignatureScores(vis)
head(vis@SigScores)
```

## Show signature scores in Seurat object

Check if cells are identical

```{r}
identical(colnames(so), rownames(vis@SigScores))
```

```{r}
so$sig.IM <- vis@SigScores[, "IM"]
so$sig.pMo <- vis@SigScores[, "pMo"]
so$sig.cMo <- vis@SigScores[, "cMo"]
```

```{r}
FeaturePlot(so, features = "sig.IM", split.by = "group")
```
```{r fig.height=4, fig.width=5}
VlnPlot(so, features = "sig.IM", cols = pal4[c(1,3,4)], pt.size = 0, idents = c("Classical Mono", "IM", "Mafb-independent"))
```
```{r}
so.sub <- subset(so, idents = c("Classical Mono", "IM", "Mafb-independent"))
dt <- data.frame(sig=so.sub$sig.IM, celltype=so.sub$cell.type2)
res.aov <- aov(sig ~ celltype, data = dt)
summary(glht(res.aov, linfct = mcp(celltype = "Tukey")))
```

```{r}
ggsave(filename = "../Figures/Vlnplot_sig_IM_in_ctrl_MafbKO.pdf", 
       width = 5, height = 4)
```

```{r fig.height=4, fig.width=5}
VlnPlot(so, features = "sig.pMo", cols = pal4[c(1,3,4)], pt.size = 0, idents = c("Classical Mono", "IM", "Mafb-independent"))

```
```{r}
dt <- data.frame(sig=so$sig.pMo, celltype=so$cell.type2)
res.aov <- aov(sig ~ celltype, data = dt)
summary(glht(res.aov, linfct = mcp(celltype = "Tukey")))
```

```{r}
ggsave(filename = "../Figures/Vlnplot_sig_pMo_in_ctrl_MafbKO.pdf", 
       width = 5, height = 4)
```

```{r}
dt <- data.frame(sig=so.sub$sig.cMo, celltype=so.sub$cell.type2)
res.aov <- aov(sig ~ celltype, data = dt)
summary(glht(res.aov, linfct = mcp(celltype = "Tukey")))
```

```{r fig.height=4, fig.width=5}
VlnPlot(so, features = "sig.cMo", cols = pal4[c(1,3,4)], pt.size = 0, idents = c("Classical Mono", "IM", "Mafb-independent"))
```
```{r}
ggsave(filename = "../Figures/Vlnplot_sig_cMo_in_ctrl_MafbKO.pdf", 
       width = 5, height = 4)
```



# Session information

R sesssion:

```{r}
sessionInfo()
```

# References
