---
subtitle: "14 - Compare with cMaf-KO IM"
# bibliography: bib.bib
link-citations: yes
date: "`r format (Sys.time(), format='%Y-%m-%d %H:%M:%S %z')`"
output: 
  #html_document
  pdf_document:
    pandoc_args:
      - '../common.yaml'
      - --listings
    includes:
      in_header: '../preamble.tex'
    toc: true
    number_sections: true
    fig_width: 7
    fig_height: 6
    fig_caption: true
    df_print: tibble
---

```{r include=FALSE}
knitr::opts_chunk$set(message = FALSE, warning = FALSE)
```

\newpage

# Description

# Load packages and data

```{r}
suppressMessages({
  library(Seurat)
  library(ggplot2)
  library(RColorBrewer)
})

so <- readRDS(file = "../12-cMAF_and_Mafb_deficient_IM/All_samples_Maf.seuratObject.Rds")
```

We will take into account the cMAF-KO and dKO samples into analysis. 

# Compare populations

## Re-process data

```{r}
so <- NormalizeData(so, verbose=FALSE)
so <- FindVariableFeatures(so, selection.method = "vst", nfeatures = 2000, verbose=FALSE) # we focus less variable genes. 
so <- ScaleData(so, features = rownames(so), verbose=FALSE)
so <- RunPCA(so, features = VariableFeatures(so), verbose=FALSE)
so <- RunTSNE(so, dims = 1:8, verbose=FALSE)
so <- RunUMAP(so, dims = 1:8, verbose=FALSE)
```


```{r message=FALSE, warning=FALSE}
so <- FindNeighbors(so, dims = 1:8, verbose = FALSE)
so <- FindClusters(so, resolution = 0.12, verbose = FALSE) 
```

```{r}
DimPlot(so, reduction = "umap")
```
```{r}
FeaturePlot(so, features = c("Mrc1", "Lyve1", "Cd72", "Mafb"))
```


```{r fig.height=16, fig.width=16}
VlnPlot(
  so, 
  features = c("Ccr2", "Plac8" ,"Ly6c2", "Spn", 
                              "Fcgr4", "Ace", "Cxcr4", 
                              "Fcgr1", "Mrc1", "Cd63", "Apoe", 
                              "H2-DMb2", "Cd74", "Cd209a", 
                              "Ccnd1","Mki67", "Top2a", "Ccl2", 
                              "Irf7", "Cxcl10", "Ifit1", "percent.mt", "Ifi47", "Itgax") , 
        ncol = 4)
```

cluster 0: Mafb-KO
Cluster 1: IM
Cluster 2: Patrolling Monocytes
Cluster 3: Classical Monocytes
Cluster 4: Unknown

```{r}
so$cell.type2 <- factor(Idents(so), labels = c("Mafb-independent", "IM", "Patrolling Mono", "Classical Mono", "Unknown"))
so$cell.type2 <- factor(so$cell.type2, levels = c("Classical Mono", "Patrolling Mono", "IM", "Mafb-independent", "Unknown"))
Idents(so) <- "cell.type2"
```

```{r fig.height=5, fig.width=5}
pal4 <- c(
  "#A6CEE3", # cMo
  "#1F78B4", # pMo
  "#33A02C", # CD206 IM
   "#FF7F00", # Mafb- neo
  "black" # unknown
)
DimPlot(so, cols = pal4, split.by = "group", ncol = 2
) + NoLegend()
```

```{r}
ggsave(filename = "../Figures/UMAPplot_All_samplesMaf_separate_2colunns.pdf", width = 5, height = 5)
```

# Identify the CD206+ and CD206- IMs

```{r}
ims <- subset(so, idents = "IM")

ims <- NormalizeData(ims, verbose=FALSE)
ims <- FindVariableFeatures(ims, selection.method = "vst", nfeatures = 2000, verbose=FALSE) # we focus less variable genes. 
ims <- ScaleData(ims, features = rownames(ims), verbose=FALSE)
ims <- RunPCA(ims, features = VariableFeatures(ims), verbose=FALSE)
ims <- RunTSNE(ims, dims = 1:8, verbose=FALSE)
ims <- RunUMAP(ims, dims = 1:8, verbose=FALSE)
ims <- FindNeighbors(ims, dims = 1:8, verbose = FALSE)
ims <- FindClusters(ims, resolution = 0.15, verbose = T) 
DimPlot(ims)
```

```{r}
VlnPlot(ims, features = c("Mrc1", "Lyve1", "Cd72", "Cd74"))
```

cluster 0: CD206+ IM
cluster 1: CD206- IM

```{r}
so$cell.type3 <- as.character(so$cell.type2)
so$cell.type3[WhichCells(ims, idents = "0")] <- "CD206+ IM"
so$cell.type3[WhichCells(ims, idents = "1")] <- "CD206- IM"
so$cell.type3 <- factor(so$cell.type3, levels = c("Classical Mono", "Patrolling Mono", "CD206+ IM", "CD206- IM", "Mafb-independent", "Unknown"))
Idents(so) <- "cell.type3"
```

```{r fig.height=5, fig.width=5}
pal4 <- c(
  "#A6CEE3", # cMo
  "#1F78B4", # pMo
  "#33A02C", # CD206 IM
  "#B2DF8A", # MHCII IM
  "#FF7F00", # Mafb- neo
  "black" # unknown
)
DimPlot(so, cols = pal4, split.by = "group", ncol = 2
) + NoLegend()
```

```{r}
ggsave(filename = "../Figures/UMAPplot_All_samplesMaf_with_IMsubsets_separate_2colunns.pdf", width = 5, height = 5)
```

# Focus on the cMAF-KO vs Control

```{r}
so <- subset(so, subset = group == c("HT5-Control", "HT6-cMAF-KO"))
```

Plot cell in colors but only for one of two samples

```{r}
so$cell.type.control <- as.character(so$cell.type3)
so$cell.type.control[WhichCells(so, expression = group == "HT6-cMAF-KO")] <- "ZZ"
so$cell.type.control <- factor(so$cell.type.control, levels = c("Classical Mono", "Patrolling Mono", "CD206+ IM", "CD206- IM", "Mafb-independent", "Unknown", "ZZ"))
Idents(so) <- "cell.type.control"
```

```{r fig.height=5, fig.width=7}
pal4.control <- c(
  "black", # unknown
  "#F1F1F1", # grey for another sample
  "#B2DF8A", # MHCII IM
  "#33A02C", # CD206 IM
  "#FF7F00", # Mafb- neo
    "#A6CEE3", # cMo
  "#1F78B4" # pMo
)
```


```{r fig.height=5, fig.width=5}
DimPlot(so, cols = pal4.control, group.by = "cell.type.control", pt.size = 2,
order = c("IM", "Patrolling Mono", "Classical Mono", "Mafb-independent", "CD206+ IM", "CD206- IM", "ZZ")
) + NoLegend()
```

```{r}
ggsave(filename = "../Figures/UMAPplot_Ctl_(vscMafKO)_no_legend.pdf", 
       width = 5, height = 5)
```

Plot Mafb-KO: 

```{r}
so$cell.type.mafbko <- as.character(so$cell.type2)
so$cell.type.mafbko[WhichCells(so, expression = group == "HT5-Control")] <- "ZZ"
so$cell.type.mafbko <- factor(so$cell.type.mafbko, levels = c("Classical Mono", "Patrolling Mono", "CD206+ IM", "CD206- IM", "Mafb-independent", "Unknown", "ZZ"))
```

```{r fig.height=5, fig.width=5}
DimPlot(so, cols = pal4.control, group.by = "cell.type.control", pt.size = 2,
order = c("IM", "Patrolling Mono", "Classical Mono", "Mafb-independent", "CD206+ IM", "CD206- IM", "ZZ")
) + NoLegend()
```

```{r}
ggsave(filename = "../Figures/UMAPplot_cMafKO_(vsControl)_no_legend.pdf", 
       width = 5, height = 5)
```

See population frequencies:

```{r fig.width=3, fig.height=4}
source("../R/SeuratFreqTable.R")
freq.celltype.list <- list(
  `Control` = Seurat2CellFreqTable(subset(so, subset = group == "HT5-Control"), slotName = "cell.type3"),  
  `MAFb-KO` = Seurat2CellFreqTable(subset(so, subset = group == "HT6-cMAF-KO"), slotName = "cell.type3") 
)

source("../R/barChart.R")
barChart(freq.celltype.list) + labs(fill = "Cluster") + scale_fill_manual(values = pal4) + theme(axis.text.x = element_text(angle = 90))
```

```{r}
ggsave(filename = "../Figures/Barplot_Ctl_cMafKO_population_frequency.pdf", 
       width = 3, height = 4)
```

Show CD206+ IM markers


```{r fig.height=4, fig.width=5}
FeaturePlot(so, features = "Mrc1", cols = c("lightgray", "red"), min.cutoff = 0.5)
```

```{r}
ggsave(filename = "../Figures/VlnPlot_Mrc1_in_cMafKO_Control.pdf", width = 5, height = 4)
```

```{r fig.height=4, fig.width=5}
FeaturePlot(so, features = "Lyve1", cols = c("lightgray", "red"))
```

```{r}
ggsave(filename = "../Figures/VlnPlot_Lyve1_in_cMafKO_control.pdf", width = 5, height = 4)
```

```{r fig.height=8, fig.width=8}
FeaturePlot(so, features = c("Mrc1", "Lyve1"), cols = c("lightgray", "red"), split.by = "group")
```
```{r}
ggsave(filename = "../Figures/Featureplot_Mrc1_Lyve1_in_cMafKO_control.pdf", height = 8, width = 8)
```

# Proliferation comparison

```{r fig.width=3, fig.height=4}
freq.celltype.list <- list(
  `Control` = Seurat2CellFreqTable(subset(so, subset = group == "HT5-Control"), slotName = "Phase"),  
  `MAFb-KO` = Seurat2CellFreqTable(subset(so, subset = group == "HT6-cMAF-KO"), slotName = "Phase") 
)

barChart(freq.celltype.list) + labs(fill = "Cluster") +  theme(axis.text.x = element_text(angle = 90))
```

```{r fig.width=9, fig.height=4}
DimPlot(so, group.by = "Phase", split.by = "group")
```

# Comparison between cMAF-deficient IMs population and control IMs

Let's focus on the Mafb-deficient population in Mafb-deficient sample.

```{r}
ims <- subset(so, subset = cell.type2 == "IM")
Idents(ims) <- "group"
```

## DE genes between Mafb- neo and IM population

```{r fig.width=8, fig.height=10}
library(dplyr)
cMafKO_vs_IM <- FindMarkers(ims, 
                          ident.1 = "HT6-cMAF-KO", 
                          ident.2 = "HT5-Control", 
                          logfc.threshold = 0,
                          verbose = FALSE)


# keep only adj p value < 0.05 and logFC > 0.5 as significant markers. 
cMafKO_vs_IM.markers <- cMafKO_vs_IM[cMafKO_vs_IM$p_val_adj < 0.05 & abs(cMafKO_vs_IM$avg_log2FC) > 0.5, ]
cMafKO_vs_IM.markers <- cMafKO_vs_IM.markers[order(cMafKO_vs_IM.markers$avg_log2FC, decreasing = TRUE), ]
nrow(cMafKO_vs_IM.markers)
```

```{r}
write.csv(cMafKO_vs_IM.markers ,file = "./cMAFKO_vs_IM.DEgenes.results.csv", quote = FALSE)
```


Show in vlnplot

```{r fig.height=8, fig.width=8}
p <- VlnPlot(ims, features = c(rownames(cMafKO_vs_IM.markers)), cols = c("#33A02C", "#FF7F00"), ncol = 4, pt.size = 0)
p
```
Show vlnplot without label: 

```{r fig.height=7, fig.width=8}
p & theme(axis.title.x=element_blank(),
        axis.text.x=element_blank(),
        axis.ticks.x=element_blank())
```
```{r}
ggsave(filename = "../Figures/Vlnplot_DE_genes_in_Ctrl_cMafKO.pdf", 
       width = 8, height = 8)
```

```{r fig.height=6, fig.width=6}
FeaturePlot(ims, features = c("Maf", "Folr2"), split.by = "group", cols = c("lightgray", "red"))

```


### Volcano plot of DE genes

```{r}
suppressMessages({
  library(dplyr)
  library(ggrepel)
  })
```


Let's set a threshold of log2FC and p_val_adj and plot them all: 

```{r fig.height=5, fig.width=6}
threshold.log2fc <- 0.5
threshold.adjp <- 0.05
cMafKO_vs_IM.volcano = mutate(cMafKO_vs_IM, 
 Sig=ifelse((abs(cMafKO_vs_IM$avg_log2FC) > threshold.log2fc)&(cMafKO_vs_IM$p_val_adj < threshold.adjp), "Sig", "n.s."))
# add two colors to 2 sig lists
cMafKO_vs_IM.volcano$Sig [ cMafKO_vs_IM.volcano$avg_log2FC < -threshold.log2fc & cMafKO_vs_IM.volcano$p_val_adj < threshold.adjp ] <- "1"

cMafKO_vs_IM.volcano$Sig [ cMafKO_vs_IM.volcano$avg_log2FC > threshold.log2fc & cMafKO_vs_IM.volcano$p_val_adj < threshold.adjp ] <- "2"

cMafKO_vs_IM.volcano$Gene <- rownames(cMafKO_vs_IM.volcano)
Gene.to.show.ValcanoPlot <- rownames(cMafKO_vs_IM.volcano[cMafKO_vs_IM.volcano$Sig != "n.s.", ])

p <- ggplot(cMafKO_vs_IM.volcano, aes(avg_log2FC, -log10(p_val_adj))) + geom_point(aes(col=Sig)) + scale_color_manual(values=c( `1`="red", `n.s.`="grey", `2`="blue")) 

# set axis lim: 
axis.lim <- max(abs(cMafKO_vs_IM.volcano$avg_log2FC)) + 1

p + geom_text_repel(data=filter(cMafKO_vs_IM.volcano, Gene %in% Gene.to.show.ValcanoPlot), size = 8, aes(label=Gene, fontface = "italic"), box.padding = 1) + xlim(c(-axis.lim, axis.lim)) + theme_classic() + theme(panel.border = element_rect(colour = "black", fill = NA, size = 1)) + geom_hline(yintercept = -log10(threshold.adjp), linetype='dashed', col = 'grey') + geom_vline(xintercept = c(-threshold.log2fc, threshold.log2fc),  linetype='dashed', col = 'grey') + ggtitle(paste("Log2FC > ", threshold.log2fc, "; p_val_adj <", threshold.adjp))

write.csv(cMafKO_vs_IM.volcano[Gene.to.show.ValcanoPlot, ] %>% arrange(., desc(avg_log2FC)) , 
          file = paste("./Mafb-deficient_vs_IM.DEgenes.Log2FC", threshold.log2fc, ".adjPval", threshold.adjp, ".results.csv", sep = "") ) 

```
```{r}
ggsave(filename = "../Figures/VolcanoPlot_DE_IM_ctrol_vs_cMafKO.pdf", 
       width = 6, height = 5)
```


  
# Session information

R sesssion:

```{r}
sessionInfo()
```

# References
