---
subtitle: "14 - Compare with cMaf-KO IM"
# bibliography: bib.bib
link-citations: yes
date: "`r format (Sys.time(), format='%Y-%m-%d %H:%M:%S %z')`"
output: 
  #html_document
  pdf_document:
    pandoc_args:
      - '../common.yaml'
      - --listings
    includes:
      in_header: '../preamble.tex'
    toc: true
    number_sections: true
    fig_width: 7
    fig_height: 6
    fig_caption: true
    df_print: tibble
---

```{r include=FALSE}
knitr::opts_chunk$set(message = FALSE, warning = FALSE)
```

\newpage

# Description

# Load packages and data

```{r}
suppressMessages({
  library(Seurat)
  library(ggplot2)
})

results <- readRDS(file = "../12-cMAF_and_Mafb_deficient_IM/All_samples_Maf.seuratObject.Rds")

# we will work on control and Mafb-KO samples, so subset: 
so <- subset(results, subset = group == c("HT5-Control", "HT6-cMAF-KO"))
```

# Compare populations

```{r, fig.height=4, fig.width=6}
pal3 <- c(
  "#A6CEE3", # cMo
  "#1F78B4", # pMo
  "#FF7F00", # Intermediate
  "#B2DF8A", # MHCII IM
  "#33A02C", # CD206 IM
  "#CAB2D6", # Mafb- neo
  "#ededed"# Unknown
)
```

As cMAF-KO has subtle difference from control, we re analyze expression data with only this two groups to focus on the difference. 

## Re-process data

```{r}
so <- NormalizeData(so, verbose=FALSE)
so <- FindVariableFeatures(so, selection.method = "vst", nfeatures = 1000, verbose=FALSE) # we focus less variable genes. 
so <- ScaleData(so, features = rownames(so), verbose=FALSE)
so <- RunPCA(so, features = VariableFeatures(so), verbose=FALSE)
so <- RunTSNE(so, dims = 1:5, verbose=FALSE)
so <- RunUMAP(so, dims = 1:5, verbose=FALSE)
```






```{r message=FALSE, warning=FALSE}
so <- FindNeighbors(so, dims = 1:5, verbose = FALSE)
so <- FindClusters(so, resolution = 0.2, verbose = FALSE) 
```

```{r fig.height=16, fig.width=16}
VlnPlot(
  so, 
  features = c("Ccr2", "Plac8" ,"Ly6c2", "Spn", 
                              "Fcgr4", "Ace", "Cxcr4", 
                              "Fcgr1", "Mrc1", "Cd63", "Apoe", 
                              "H2-DMb2", "Cd74", "Cd209a", 
                              "Ccnd1","Mki67", "Top2a", "Ccl2", 
                              "Irf7", "Cxcl10", "Ifit1", "percent.mt", "Ifi47", "Itgax") , 
        ncol = 4, cols = brewer.pal(5, "Paired"))
```

cluster 0: Intermediate
Cluster 1: CD206+ IM
Cluster 2: CD206-/MHCII IM
Cluster 3: Classical Monocytes
Cluster 4: Patrolling Monocytes

```{r}
so$cell.type3 <- factor(Idents(so), labels = c("Intermediate", "CD206+ IMs", "CD206- IMs", "Classical Mono", "Patrolling Mono"))
so$cell.type3 <- factor(so$cell.type3, levels = c("Classical Mono", "Patrolling Mono", "Intermediate", "CD206- IMs", "CD206+ IMs"))
Idents(so) <- "cell.type3"
DimPlot(so, cols = pal3[1:5])
```


```{r}
ggsave(filename = "../Figures/UMAPplot_Ctl_cMafKO_with_legend.pdf", 
       width = 6, height = 4)
```


```{r}
DimPlot(so, cols = pal3[1:5], split.by = "group")
```


```{r include=FALSE, fig.height=4, fig.width=5}
DimPlot(so, cols = pal3) + NoLegend()
ggsave(filename = "../Figures/UMAPplot_Ctl_cMafKO_no_legend.pdf", 
       width = 5, height = 4)

DimPlot(subset(so, subset = group == "HT6-cMAF-KO"), 
        cols = pal3) + NoLegend()
ggsave(filename = "../Figures/UMAPplot_cMafKO_no_legend.pdf", 
       width = 5, height = 4)

DimPlot(so, group.by = "orig.ident",
        cols = "#f5f5f5") + NoLegend()
ggsave(filename = "../Figures/UMAPplot_Ctl_MafbKO_grey.pdf", 
       width = 5, height = 4)
```

See population frequencies:

```{r fig.width=3, fig.height=4}
source("../R/SeuratFreqTable.R")
freq.celltype.list <- list(
  `Control` = Seurat2CellFreqTable(subset(so, subset = group == "HT5-Control"), slotName = "cell.type3"),  
  `cMAF-KO` = Seurat2CellFreqTable(subset(so, subset = group == "HT6-cMAF-KO"), slotName = "cell.type3") 
)

source("../R/barChart.R")
barChart(freq.celltype.list) + labs(fill = "Cluster") + scale_fill_manual(values = pal3) + theme(axis.text.x = element_text(angle = 90))
```

```{r}
ggsave(filename = "../Figures/Barplot_Ctl_MafbKO_population_frequency.pdf", 
       width = 3, height = 4)
```

# Proliferation comparison

```{r fig.width=3, fig.height=4}
freq.celltype.list <- list(
  `Control` = Seurat2CellFreqTable(subset(so, subset = group == "HT5-Control"), slotName = "Phase"),  
  `MAFb-KO` = Seurat2CellFreqTable(subset(so, subset = group == "HT6-cMAF-KO"), slotName = "Phase") 
)

barChart(freq.celltype.list) + labs(fill = "Cluster") +  theme(axis.text.x = element_text(angle = 90))
```

```{r fig.width=9, fig.height=4}
DimPlot(so, group.by = "Phase", split.by = "group")
```

```{r}
VlnPlot(so, features = "G2M.Score", split.by = "group", cols = pal3, split.plot = TRUE)
```

# Comparison between cMaf-KO and IMs

Let's focus on the cMaf-deficient population in Mafb-deficient sample.

```{r}
IM <- subset(so, subset = cell.type3 %in% c("CD206- IMs", "CD206+ IMs"))
```

## DE genes between cMaf-KO and control IM population

```{r fig.width=8, fig.height=10}
library(dplyr)
Idents(IM) <- "group"
cMafKO_vs_Ctl <- FindMarkers(IM, 
                          ident.1 = "HT6-cMAF-KO", 
                          ident.2 = "HT5-Control", 
                          logfc.threshold = 0,
                          verbose = FALSE)


# keep only adj p value < 0.05 and logFC > 0.5 as significant markers. 
cMafKO_vs_Ctl.markers <- cMafKO_vs_Ctl[cMafKO_vs_Ctl$p_val_adj < 0.05 & abs(cMafKO_vs_Ctl$avg_log2FC) > 0.5, ]
cMafKO_vs_Ctl.markers <- cMafKO_vs_Ctl.markers[order(cMafKO_vs_Ctl.markers$avg_log2FC, decreasing = TRUE), ]
nrow(cMafKO_vs_Ctl.markers)
```

```{r}
write.csv(cMafKO_vs_Ctl.markers ,file = "./Mafb-deficient_vs_IM.DEgenes.results.csv", quote = FALSE)
```


Let's show the top 20 of each side: 

```{r}
cMafKO_vs_Ctl.markers.top20 <- cMafKO_vs_Ctl.markers[c(1:20, (nrow(cMafKO_vs_Ctl.markers)-19):(nrow(cMafKO_vs_Ctl.markers))), ]
```


```{r message=FALSE}
library(ComplexHeatmap)
mat <- GetAssayData(neo_IM)[rownames(cMafKO_vs_Ctl.markers.top20), ]
mat.scale <- t(scale(t(as.matrix(mat))))
```

```{r fig.height=5, fig.width=6}
hp <- Heatmap(mat.scale, show_row_names = TRUE, show_column_names = FALSE, 
        row_names_gp = gpar ( fontsize = 7), 
        column_split = factor(neo_IM$cell.type2), 
        km = 2)
hp <- draw(hp)
```


```{r}
pdf(file = "../Figures/Heatmap_IM_vs_Mafb-neo.pdf", width = 6, height = 5)
hp
dev.off()
```

Show in vlnplot

```{r fig.height=8, fig.width=8}
VlnPlot(neo_IM, features = c("Mafb", "Mrc1", "Mertk", "Cd38", "Adgre1", "C1qa", "Folr2", "Cd163"), group.by = "cell.type2", cols = c("#33A02C",  "#CAB2D6"))
```


## GO enrichment analysis with DE genes

```{r}
suppressMessages(library(clusterProfiler))
source("../R/entrez2symbol.R")
source("../R/replaceEntrezID.R")
```

### GO enrichment analysis of up-regulated DE genes in Mafb-deficient population

```{r message=FALSE}
DE.MafbKO <- cMafKO_vs_Ctl.markers[cMafKO_vs_Ctl.markers$avg_log2FC > 0, ]
symb <- rownames(DE.MafbKO)
de_entrez <- bitr( geneID = symb, fromType = "SYMBOL", toType = "ENTREZID", OrgDb = "org.Mm.eg.db", drop = TRUE ) $ ENTREZID
result.enrichGO <- enrichGO(de_entrez, OrgDb = "org.Mm.eg.db", ont = "BP")
result.enrichGO <- replaceEntrezID(result.enrichGO, organism = "mmu")
write.csv(result.enrichGO@result, file = "./Results_enrichment/enrichGO_DE_Mafb_KO_vsIM.csv")
result.enrichGO@result
```

### GO enrichment analysis of up-regulated DE genes in IMs

```{r message=FALSE}
DE.IM <- cMafKO_vs_Ctl.markers[cMafKO_vs_Ctl.markers$avg_log2FC < 0, ]
symb <- rownames(DE.IM)
de_entrez <- bitr( geneID = symb, fromType = "SYMBOL", toType = "ENTREZID", OrgDb = "org.Mm.eg.db", drop = TRUE ) $ ENTREZID
result.enrichGO <- enrichGO(de_entrez, OrgDb = "org.Mm.eg.db", ont = "BP")
result.enrichGO <- replaceEntrezID(result.enrichGO, organism = "mmu")
write.csv(result.enrichGO@result, file = "./Results_enrichment/enrichGO_DE_IM_vsMafbKO.csv")
result.enrichGO@result
```

### Volcano plot of DE genes

```{r}
suppressMessages({
  library(dplyr)
  library(ggrepel)
  })

cMafKO_vs_Ctl.volcano = mutate(cMafKO_vs_Ctl, 
 Sig=ifelse((abs(cMafKO_vs_Ctl$avg_log2FC)>0.5)&(cMafKO_vs_Ctl$p_val_adj < 0.05), "Sig", "n.s."))
# res2 = mutate(res2, 
#              Sig=ifelse((abs(res2$avg_logFC)>1)&(res2$p_val_adj < 0.01), "Sig", "n.s."))


cMafKO_vs_Ctl.volcano$Gene <- rownames(cMafKO_vs_Ctl.volcano)
Gene.to.show.ValcanoPlot <- c("C1qa", "C1qb", "C1qc", "Maf", "Cd209", "Mafb")
```


```{r fig.height=5, fig.width=6}
p <- ggplot(cMafKO_vs_Ctl.volcano, aes(avg_log2FC, -log10(p_val_adj))) + geom_point(aes(col=Sig)) + scale_color_manual(values=c( "gray", "red")) 

p + geom_text_repel(data=filter(cMafKO_vs_Ctl.volcano, Gene %in% Gene.to.show.ValcanoPlot), aes(label=Gene), box.padding = 1) + xlim(c(-5, 5)) + theme_classic() + theme(panel.border = element_rect(colour = "black", fill = NA, size = 1)) + geom_hline(yintercept = -log10(0.01), linetype='dashed', col = 'grey') + geom_vline(xintercept = c(-0.5, 0.5),  linetype='dashed', col = 'grey')

```

```{r}
ggsave(filename = "../Figures/VolcanoPlot_Mafb-KO_vs_IM_DE_genes.pdf", height = 5, width = 6)
```

# Session information

R sesssion:

```{r}
sessionInfo()
```

# References
