---
subtitle: "10-DE genes across pseudotime"
# bibliography: bib.bib
link-citations: yes
date: "`r format (Sys.time(), format='%Y-%m-%d %H:%M:%S %z')`"
output: 
  #html_document
  pdf_document:
    pandoc_args:
      - '../common.yaml'
      - --listings
    includes:
      in_header: '../preamble.tex'
    toc: true
    number_sections: true
    fig_width: 7
    fig_height: 6
    fig_caption: true
    df_print: tibble
---

```{r include=FALSE}
knitr::opts_chunk$set(message = FALSE, warning = FALSE)
```

\newpage

# Description





# Prepare data

```{r}
suppressMessages(
{
library(Seurat)
library(ComplexHeatmap)
library(ggplot2)
library(dplyr)
library(RColorBrewer)
library(circlize)
library(monocle3)
})

mo <- readRDS(file = "../9-Monocle_analysis_and_pseudotime_estimation/Mono_to_IM.cds")

```


Show trajectory in UMAP plot: 

```{r fig.height=5, fig.width=6.5}
pal2 <- c(`Classical_Mono`="#A6CEE3", 
          `MHCII_IM`="#B2DF8A",
          `CD206_IM`="#33A02C", 
          `Transit` = "#E31A1C")

plot_cells(cds = mo, color_cells_by = "cell.type2", 
           cell_size = 1, label_cell_groups = FALSE, 
           label_branch_points = TRUE, label_leaves = TRUE, 
           label_roots = FALSE, alpha = 1) + 
  scale_color_manual(values = pal2)
```

```{r eval=FALSE}
ggsave(filename = "../Figures/UMAPplot_trajectory.pdf", width = 6.5, height = 5)
```


# DE gene expression across IM-differentiation

[DE genes across pseudotime of IM differentiation](#de-genes)
## Across pseudotime of IM differentiation

```{r echo=FALSE, results='hide', message = FALSE}
modulated_genes <- graph_test(mo, neighbor_graph = "principal_graph", cores = 4)
genes <- row.names(subset(modulated_genes, q_value == 0 & morans_I > 0.25))
length(genes)
```

Prepare matrix with z-scores, smoothened and scaled data across pseudotime for heatmap. 

```{r}
pt.matrix <- exprs(mo)[match(genes,rownames(rowData(mo))),order(pseudotime(mo))]
cellnames <- colnames(pt.matrix)
#Can also use "normalized_counts" instead of "exprs" to use various normalization methods, for example:
#normalized_counts(cds, norm_method = "log")

pt.matrix <- t(apply(pt.matrix,1,function(x){smooth.spline(x,df=3)$y}))
pt.matrix <- t(apply(pt.matrix,1,function(x){(x-mean(x))/sd(x)}))
rownames(pt.matrix) <- genes
colnames(pt.matrix) <- cellnames
```

Show DE genes in unsupervised heatmap. 

```{r fig.width=8, fig.height=10}
#K means with 6 groups
htkm <- Heatmap(
  pt.matrix,
  # use_raster = FALSE, # use FALSE to export to vector image. 
  name                         = "z-score",
  col                          = colorRamp2(seq(from=-2,to=2,length=11),rev(brewer.pal(11, "Spectral"))),
  show_row_names               = FALSE,
  show_column_names            = FALSE,
  row_names_gp                 = gpar(fontsize = 3),
  row_km = 6, 
  row_km_repeats = 31, 
  row_dend_reorder = TRUE, 
  row_title_rot                = 0,
  cluster_rows                 = TRUE,
  cluster_row_slices           = FALSE,
  cluster_columns              = FALSE, 
)

htkm <- draw(htkm)
```
*In this heatmap, the x axis is pseudotime, which represents differentiation state from monocytes (left) to IMs (right). *

## Annotate the cells associated to either differentiation of CD206+ IMs or CD206- IMs

```{r}
library(magrittr)
# Get the closest vertice for every cell
y_to_cells <-  mo@principal_graph_aux$UMAP$pr_graph_cell_proj_closest_vertex%>%as.data.frame()

y_to_cells$cells <- rownames(y_to_cells)
y_to_cells$Y <- y_to_cells$V1


# Get the root vertices
# It is the same node as above
root <- mo@principal_graph_aux$UMAP$root_pr_nodes

principalgraph <- mo@principal_graph$UMAP

# Get the other endpoints
endpoints <- names(which(igraph::degree(principalgraph ) == 1))
endpoints <- endpoints[!endpoints %in% root]

# For each endpoint
cellWeights <- lapply(endpoints, function(endpoint) {
  # We find the path between the endpoint and the root
  path <- igraph::shortest_paths(principalgraph, root, endpoint)$vpath[[1]]
  path <- as.character(path)
  # We find the cells that map along that path
  df <- y_to_cells[y_to_cells$Y %in% path, ]
  df <- data.frame(weights = as.numeric(colnames(mo) %in% df$cells))
  colnames(df) <- endpoint
  return(df)
  }) %>% do.call(what = 'cbind', args = .) %>%
    as.matrix()
rownames(cellWeights) <- colnames(mo)
colnames(cellWeights) <- c("CD206_IM branch", "MHCII_IM branch")
pseudotime <- matrix(mo@principal_graph_aux$UMAP$pseudotime, ncol = ncol(cellWeights),
                     nrow = ncol(mo), byrow = FALSE)
rownames(pseudotime) <- colnames(mo)
```

# TradeSeq analysis for the differentiation of monocytes to either of IM subsets 

## Construct sce object for TradeSeq


```{r}
suppressMessages(library(tradeSeq))
```

```{r eval=FALSE}
# this step is VERY time consuming
sce <- fitGAM(counts = mo@assays@data$counts, 
              pseudotime = pseudotime, 
              cellWeights = cellWeights)

saveRDS(sce, file = "./sce.4339cells.newversion.Rds")
```

```{r include=FALSE}
sce <- readRDS(file = "./sce.4339cells.Rds")
```

[Between-lineage comparisons (CD206+ IM vs CD206- IM differentiation)](#wald-calc)
## Between-lineage comparisons (CD206+ IM vs CD206- IM differentiation)

Association of gene expression with pseudotime (find significant DE genes along pseudotime). 

```{r}
assoRes <- associationTest(sce)
endRes <- diffEndTest(sce)
head(assoRes)
```

Plot the most sig gene: 
```{r}
library(ggplot2)
o <- order(endRes$waldStat, decreasing = TRUE)
sigGene <- names(sce)[o[1]]
plotSmoothers(sce, counts = counts(sce), gene = sigGene
              #, curvesCol = c("#33A02C", "#B2DF8A")
              ) + ggtitle(sigGene) 
```

What's the top genes? 
```{r}
names(sce)[o[1:20]]
```

## Clustering using RSEC, clusterExperiment

> tradeSeq provides the functionality to cluster genes according to their expression pattern along the lineages with the clusterExpressionPatterns function. A number of equally spaced points for every lineage are selected to perform the clustering, and the number of points can be selected with the nPoints argument. (from `vignette("tradeSeq")`)

```{r}
library(clusterExperiment)
nPointsClus <- 20 # The number of points to use for clustering the expression patterns.. 
clusPat <- clusterExpressionPatterns(sce, 
                                     nPoints = nPointsClus,
                                     genes = genes,
                                     random.seed = 43, 
                                     beta = 0.2
                                     )
```

```{r}
clusterLabels <- primaryCluster(clusPat$rsec)
```


```{r fig.width=12, fig.height=12}
cUniq <- unique(clusterLabels) # 
cUniq <- cUniq[!cUniq == -1] # remove unclustered genes 

# cUniq <- cUniq[cUniq == -1]
#Any samples not found as part of a homogenous set of clusters at that point will be classified as unclustered (given a value of -1)

# beta: value between 0 and 1 to decide how stable clustership membership has to be before 'finding' and removing the cluster.
if (exists("p.total")) { rm(p.total)}

for (xx in cUniq) {
  cId <- which(clusterLabels == xx)
  p <- ggplot(data = data.frame(x = 1:nPointsClus,
                                y = rep(range(clusPat$yhatScaled[cId, ]),
                                        nPointsClus / 2)),
              aes(x = x, y = y)) +
    geom_point(alpha = 0) +
    labs(title = paste0("Cluster ", xx),  x = "Pseudotime", y = "Normalized expression") +
    theme_classic() +
    theme(plot.title = element_text(hjust = 0.5))
  for (ii in 1:length(cId)) {
    geneId <- rownames(clusPat$yhatScaled)[cId[ii]]
    p <- p +
      geom_line(data = data.frame(x = rep(1:nPointsClus, 2),
                                  y = clusPat$yhatScaled[geneId, ],
                                  lineage = rep(0:1, each = nPointsClus)),
                aes(col = as.character(lineage), group = lineage), lwd = 1.5)
  }
  p <- p + guides(color = "none") +
    scale_color_manual(values = c("#33A02C", "#B2DF8A"),
                       breaks = c("0", "1"))
  if (exists("p.total")) { p.total <- p.total + p} else {p.total <- p}
}
  print(p.total)
```

# Show gene expression pattern calculated by TradeSeq in heatmap

## Data preparation

Here we use the DE genes calculated in [DE genes across pseudotime of IM differentiation](#de-genes). 

```{r}
yhatSmooth <- predictSmooth(sce, gene = genes, nPoints = 100, tidy = FALSE)
yhatSmoothScaled <- t(scale(t(yhatSmooth)))
```


## Draw heatmap

```{r}
heatSmooth <- Heatmap(yhatSmoothScaled, cluster_columns = FALSE, show_row_names = FALSE, show_column_names = FALSE, row_km = 4)
heatSmooth <- draw(heatSmooth)
```
*Two IM differentiation show similar patterns but some genes (especially cluster 2 and 1) are different in CD206+ and CD206-.  *

## Annotate DE genes as CD206+/CD206- IM differentiation specific or common genes

According to the heatmap above, some of DE genes should remain unchanged (common) and half of them are specific to one of two IM differentiation. 

We use wald statistic calculated in diffEndTest to annotate the "common" genes and "specific" genes. (in [Between-lineage comparisons (CD206+ IM vs CD206- IM differentiation)](#wald-calc))

```{r}
endRes.DE <- endRes[rownames(yhatSmooth), ]
summary(endRes.DE$waldStat)
```

Let's use waldStat > 40 and logFC > 2 as cut threshold. 

```{r}
genes.changed <- rownames(filter(endRes.DE, waldStat > 70 & (logFC1_2 > 2 | logFC1_2 < -2) ))
genes.noChange <- setdiff(rownames(endRes.DE) , genes.changed)
```

Make heatmap with unchanged/common genes. 

```{r fig.width=8, fig.height=12}
heatSmooth_cd206.unchanged <- Heatmap(yhatSmoothScaled[genes.noChange, 100:1], cluster_columns = FALSE,  show_row_names = FALSE, show_column_names = FALSE, column_title = "CD206+ IM")

heatSmooth_mhcii.unchanged <- Heatmap(yhatSmoothScaled[genes.noChange, 101:200], cluster_columns = FALSE,  show_row_names = TRUE, row_names_gp = gpar(fontsize = 3), show_column_names = FALSE, column_title = "MHCII+ IM")

heatSmooth_combined.unchanged <- draw ( heatSmooth_cd206.unchanged + heatSmooth_mhcii.unchanged, column_title = "Common genes", auto_adjust = FALSE)
```

## Daw heatmap with expression patterns of unchanged/common genes in the order of pseudotime

Let's find the expression peak of each gene: 

```{r}
orderbyExpressionPeak <- function(x, # matrix
                                  decreasing = FALSE, 
                                  output.position = FALSE # if true, give relative position 0 - 1, or output order. 
                                  ) {
  indx.peak <- apply(x, 1 , which.max)
  
  if(output.position) {
    po <- indx.peak/nrow(x)
    if (! length(rownames(x)) == 0) {names(po) <- rownames(x)}
    return(po)
  } else {
    o <- order(indx.peak) 
    if (! length(rownames(x)) == 0) {names(o) <- rownames(x)[o]}
    return (o)}
}
```

Make average peak pseudotime peak for each gene: 

```{r fig.height=10, fig.width=8}
po.cd206 <- orderbyExpressionPeak(yhatSmoothScaled[genes.noChange, 1:100], output.position = TRUE)
po.mhcii <- orderbyExpressionPeak(yhatSmoothScaled[genes.noChange, 101:200], output.position = TRUE)
order.mean <- order ( ( po.cd206 + po.mhcii ) /2)

heatSmooth_cd206.unchanged.ordered <- Heatmap(yhatSmoothScaled[genes.noChange, 100:1], cluster_columns = FALSE,  show_row_names = FALSE, show_column_names = FALSE, row_order = order.mean,  column_title = "CD206+ IM")
```



```{r fig.height=10, fig.width=8}
heatSmooth_combined.unchanged.ordered <- draw ( heatSmooth_cd206.unchanged.ordered + heatSmooth_mhcii.unchanged, column_title = "Common genes", auto_adjust = FALSE)
```

```{r}
pdf(file = "../Figures/Heatmap_common_genes_IMs_diff_across_pseudotime.pdf", width = 8, height = 10)
heatSmooth_combined.unchanged.ordered
dev.off()
```


## Make with changed/specific genes

```{r fig.height=8, fig.width=8}
heatSmooth_cd206.changed <- Heatmap(yhatSmoothScaled[genes.changed, 100:1], cluster_columns = FALSE,  show_row_names = FALSE, cluster_rows = hclust(dist(yhatSmoothScaled[genes.changed, ])), show_column_names = FALSE, column_title = "CD206+ IM")

heatSmooth_mhcii.changed <- Heatmap(yhatSmoothScaled[genes.changed, 101:200], cluster_columns = FALSE,  show_row_names = TRUE,row_names_gp = gpar(fontsize = 5), show_column_names = FALSE, column_title = "MHCII+ IM")

heatSmooth_combined.changed <- draw ( heatSmooth_cd206.changed + heatSmooth_mhcii.changed, column_title = "Specific genes", split = 2)
```

```{r}
pdf(file = "../Figures/Heatmap_specific_CD206_IM_genes_IMs_diff_across_pseudotime.pdf", width = 8, height = 8)
heatSmooth_combined.changed
dev.off()
```


```{r}
geneSymbol.changed.mhcii <- read.csv("./geneSymbol_changed_mhcii_IM.csv", col.names = FALSE)[[1]]
geneSymbol.changed.cd206 <- read.csv("./geneSymbol_changed_cd206_IM.csv", col.names = FALSE)[[1]]
```

MHCII IM high genes: 

```{r}
geneSymbol.changed.mhcii
```

```{r}
geneSymbol.changed.cd206
```

```{r}
write.csv(geneSymbol.changed.cd206, file = "./geneSymbol_changed_cd206_IM.csv", quote = FALSE, row.names = FALSE)
write.csv(geneSymbol.changed.mhcii, file = "./geneSymbol_changed_mhcii_IM.csv", quote = FALSE, row.names = FALSE)

```


## GO enrichment on the changed/specific genes

```{r}
suppressMessages(library(clusterProfiler))
source("../R/entrez2symbol.R")
source("../R/replaceEntrezID.R")
```

### GO enrichment on MHCII IM

```{r message=FALSE}
symb <- geneSymbol.changed.mhcii
de_entrez <- bitr( geneID = symb, fromType = "SYMBOL", toType = "ENTREZID", OrgDb = "org.Mm.eg.db", drop = TRUE ) $ ENTREZID
result.enrichGO <- enrichGO(de_entrez, OrgDb = "org.Mm.eg.db", ont = "BP")
result.enrichGO <- replaceEntrezID(result.enrichGO, organism = "mmu")
write.csv(result.enrichGO@result, file = "./Results_enrichment/enrichGO_MHCII_IM_specific_genes.csv")
result.enrichGO@result
```



### GO enrichment on CD206+ IM

```{r message=FALSE}
symb <- geneSymbol.changed.cd206
de_entrez <- bitr( geneID = symb, fromType = "SYMBOL", toType = "ENTREZID", OrgDb = "org.Mm.eg.db", drop = TRUE ) $ ENTREZID
result.enrichGO <- enrichGO(de_entrez, OrgDb = "org.Mm.eg.db", ont = "BP")
result.enrichGO <- replaceEntrezID(result.enrichGO, organism = "mmu")
write.csv(result.enrichGO@result, file = "./Results_enrichment/enrichGO_CD206_IM_specific_genes.csv")
result.enrichGO@result
```

# Functionality analysis of DE genes across pseudotime (common genes)

```{r}
common.genes <- heatSmooth_cd206.unchanged.ordered@row_names_param$labels[heatSmooth_cd206.unchanged.ordered@row_order]
```

## Manually classify genes by expression timing peak

Class-1 genes are the genes expressed in monocytes but turned off in the very early phase: 

```{r}
genes.class1 <- common.genes[1:which(common.genes == "Gm21188")]
length(genes.class1)
```


Class-2 genes are the genes up-regulated in early phase of differentiation and turned off during transit phase. 

```{r}
genes.class2 <- common.genes[(which(common.genes == "Gm21188")+1) : which(common.genes == "Diaph3")]
length(genes.class2)
```

Class-3 genes are the late upregulated genes during IM differentiation. 

```{r}
genes.class3 <- common.genes[(which(common.genes == "Diaph3")+1) : length(common.genes)]
length(genes.class3)
```

Save gene lists: 

```{r}
write.csv(genes.class1, file = "./common_genes_class1.csv", quote = FALSE)
write.csv(genes.class2, file = "./common_genes_class2.csv", quote = FALSE)
write.csv(genes.class3, file = "./common_genes_class3.csv", quote = FALSE)
```


## GO/KEGG enrichment analysis with 3 classes of common genes


### KEGG enrichment for common genes class 1

```{r message=FALSE}
symb <- genes.class1
de_entrez <- bitr( geneID = symb, fromType = "SYMBOL", toType = "ENTREZID", OrgDb = "org.Mm.eg.db", drop = TRUE ) $ ENTREZID
result.enrichKEGG <- enrichKEGG(de_entrez, organism = "mmu", keyType = "ncbi-geneid")
result.enrichKEGG <- replaceEntrezID(result.enrichKEGG, organism = "mmu")
write.csv(result.enrichKEGG@result, file = "./Results_enrichment/enrichKEGG_common_genes_class1.csv")
result.enrichKEGG@result
```

### GO enrichment for common genes class 1 

```{r message=FALSE}
result.enrichGO <- enrichGO(de_entrez, OrgDb = "org.Mm.eg.db", ont = "BP")
result.enrichGO <- replaceEntrezID(result.enrichGO, organism = "mmu")
write.csv(result.enrichGO@result, file = "./Results_enrichment/enrichGO_common_genes_class1.csv")
result.enrichGO@result
```

### KEGG enrichment for common genes class 2

```{r message=FALSE}
symb <- genes.class2
de_entrez <- bitr( geneID = symb, fromType = "SYMBOL", toType = "ENTREZID", OrgDb = "org.Mm.eg.db", drop = TRUE ) $ ENTREZID
result.enrichKEGG <- enrichKEGG(de_entrez, organism = "mmu", keyType = "ncbi-geneid")
result.enrichKEGG <- replaceEntrezID(result.enrichKEGG, organism = "mmu")
write.csv(result.enrichKEGG@result, file = "./Results_enrichment/enrichKEGG_common_genes_class2.csv")
result.enrichKEGG@result
```

### GO enrichment for common genes class 2 

```{r message=FALSE}
result.enrichGO <- enrichGO(de_entrez, OrgDb = "org.Mm.eg.db", ont = "BP")
result.enrichGO <- replaceEntrezID(result.enrichGO, organism = "mmu")
write.csv(result.enrichGO@result, file = "./Results_enrichment/enrichGO_common_genes_class2.csv")
result.enrichGO@result
```


### KEGG enrichment for common genes class 3

```{r message=FALSE}
symb <- genes.class3
de_entrez <- bitr( geneID = symb, fromType = "SYMBOL", toType = "ENTREZID", OrgDb = "org.Mm.eg.db", drop = TRUE ) $ ENTREZID
result.enrichKEGG <- enrichKEGG(de_entrez, organism = "mmu", keyType = "ncbi-geneid")
result.enrichKEGG <- replaceEntrezID(result.enrichKEGG, organism = "mmu")
write.csv(result.enrichKEGG@result, file = "./Results_enrichment/enrichKEGG_common_genes_class3.csv")
result.enrichKEGG@result
```

### GO enrichment for common genes class 3 

```{r message=FALSE}
result.enrichGO <- enrichGO(de_entrez, OrgDb = "org.Mm.eg.db", ont = "BP")
result.enrichGO <- replaceEntrezID(result.enrichGO, organism = "mmu")
write.csv(result.enrichGO@result, file = "./Results_enrichment/enrichGO_common_genes_class3.csv")
result.enrichGO@result
```

### GO enrichment for MHCII IM specific genes + common genes class 3 

As the specific genes have limited number. We try add the class 3 genes, which are also terminated differentiation genes. But we have to pay attention that the class 3 gene number is huge compare to the specific genes, so enrichment results will be diluted. 

```{r message=FALSE}
symb <- c(geneSymbol.changed.mhcii, genes.class3)
de_entrez <- bitr( geneID = symb, fromType = "SYMBOL", toType = "ENTREZID", OrgDb = "org.Mm.eg.db", drop = TRUE ) $ ENTREZID
result.enrichGO <- enrichGO(de_entrez, OrgDb = "org.Mm.eg.db", ont = "BP")
result.enrichGO <- replaceEntrezID(result.enrichGO, organism = "mmu")
write.csv(result.enrichGO@result, file = "./Results_enrichment/enrichGO_MHCIISpecific_plus_Class3_common.csv")
result.enrichGO@result
```

### GO enrichment for CD206 IM specific genes + common genes class 3 

```{r message=FALSE}
symb <- c(geneSymbol.changed.cd206, genes.class3)
de_entrez <- bitr( geneID = symb, fromType = "SYMBOL", toType = "ENTREZID", OrgDb = "org.Mm.eg.db", drop = TRUE ) $ ENTREZID
result.enrichGO <- enrichGO(de_entrez, OrgDb = "org.Mm.eg.db", ont = "BP")
result.enrichGO <- replaceEntrezID(result.enrichGO, organism = "mmu")
write.csv(result.enrichGO@result, file = "./Results_enrichment/enrichGO_CD206Specific_plus_Class3_common.csv")
result.enrichGO@result
```

# Show gene expression pattern with TradeSeq results

## Class 1 common genes

```{r fig.width=4, fig.height=3, message=FALSE}
require(ggplot2)
sigGene <- "Ly6c2"
plotSmoothers(sce, counts = counts(sce), gene = sigGene) + ggtitle(sigGene) + scale_color_manual(values =c("#33A02C", "#B2DF8A"))
ggsave(filename = paste("../Figures/TradeSeq_plot_IM_lineages_", sigGene, ".pdf"), 
       width = 4, height = 3)
```

```{r fig.width=4, fig.height=3, message=FALSE}
sigGene <- "Sell"
plotSmoothers(sce, counts = counts(sce), gene = sigGene) + ggtitle(sigGene) + scale_color_manual(values =c("#33A02C", "#B2DF8A"))
```


```{r fig.width=4, fig.height=3, message=FALSE}
sigGene <- "Plac8"
plotSmoothers(sce, counts = counts(sce), gene = sigGene) + ggtitle(sigGene) + scale_color_manual(values =c("#33A02C", "#B2DF8A"))
ggsave(filename = paste("../Figures/TradeSeq_plot_IM_lineages_", sigGene, ".pdf"), 
       width = 4, height = 3)
```

## Class 2 common genes

```{r fig.width=4, fig.height=3, message=FALSE}
sigGene <- "Top2a"
plotSmoothers(sce, counts = counts(sce), gene = sigGene) + ggtitle(sigGene) + scale_color_manual(values =c("#33A02C", "#B2DF8A"))
ggsave(filename = paste("../Figures/TradeSeq_plot_IM_lineages_", sigGene, ".pdf"), 
       width = 4, height = 3)
```

```{r fig.width=4, fig.height=3, message=FALSE}
sigGene <- "Ccnb2"
plotSmoothers(sce, counts = counts(sce), gene = sigGene) + ggtitle(sigGene) + scale_color_manual(values =c("#33A02C", "#B2DF8A"))
ggsave(filename = paste("../Figures/TradeSeq_plot_IM_lineages_", sigGene, ".pdf"), 
       width = 4, height = 3)
```

```{r fig.width=4, fig.height=3, message=FALSE}
sigGene <- "Mki67"
plotSmoothers(sce, counts = counts(sce), gene = sigGene) + ggtitle(sigGene) + scale_color_manual(values =c("#33A02C", "#B2DF8A"))
ggsave(filename = paste("../Figures/TradeSeq_plot_IM_lineages_", sigGene, ".pdf"), 
       width = 4, height = 3)
```

## Class 3 common genes

```{r fig.width=4, fig.height=3, message=FALSE}
sigGene <- "Tgfbr1"
plotSmoothers(sce, counts = counts(sce), gene = sigGene) + ggtitle(sigGene) + scale_color_manual(values =c("#33A02C", "#B2DF8A"))
ggsave(filename = paste("../Figures/TradeSeq_plot_IM_lineages_", sigGene, ".pdf"), 
       width = 4, height = 3)
```

```{r fig.width=4, fig.height=3, message=FALSE}
sigGene <- "Cd163"
plotSmoothers(sce, counts = counts(sce), gene = sigGene) + ggtitle(sigGene) + scale_color_manual(values =c("#33A02C", "#B2DF8A"))
ggsave(filename = paste("../Figures/TradeSeq_plot_IM_lineages_", sigGene, ".pdf"), 
       width = 4, height = 3)
```

```{r fig.width=4, fig.height=3, message=FALSE}
sigGene <- "Creb5"
plotSmoothers(sce, counts = counts(sce), gene = sigGene) + ggtitle(sigGene) + scale_color_manual(values =c("#33A02C", "#B2DF8A"))
ggsave(filename = paste("../Figures/TradeSeq_plot_IM_lineages_", sigGene, ".pdf"), 
       width = 4, height = 3)
```

```{r fig.width=4, fig.height=3, message=FALSE}
sigGene <- "Ube2c"
plotSmoothers(sce, counts = counts(sce), gene = sigGene) + ggtitle(sigGene) + scale_color_manual(values =c("#33A02C", "#B2DF8A"))
ggsave(filename = paste("../Figures/TradeSeq_plot_IM_lineages_", sigGene, ".pdf"), 
       width = 4, height = 3)
```

```{r fig.width=4, fig.height=3, message=FALSE}
sigGene <- "Aurkb"
plotSmoothers(sce, counts = counts(sce), gene = sigGene) + ggtitle(sigGene) + scale_color_manual(values =c("#33A02C", "#B2DF8A"))
ggsave(filename = paste("../Figures/TradeSeq_plot_IM_lineages_", sigGene, ".pdf"), 
       width = 4, height = 3)
```

```{r fig.width=4, fig.height=3, message=FALSE}
sigGene <- "Racgap1"
plotSmoothers(sce, counts = counts(sce), gene = sigGene) + ggtitle(sigGene) + scale_color_manual(values =c("#33A02C", "#B2DF8A"))
ggsave(filename = paste("../Figures/TradeSeq_plot_IM_lineages_", sigGene, ".pdf"), 
       width = 4, height = 3)
```

```{r fig.width=4, fig.height=3, message=FALSE}
sigGene <- "Ube2c"
plotSmoothers(sce, counts = counts(sce), gene = sigGene) + ggtitle(sigGene) + scale_color_manual(values =c("#33A02C", "#B2DF8A"))
ggsave(filename = paste("../Figures/TradeSeq_plot_IM_lineages_", sigGene, ".pdf"), 
       width = 4, height = 3)
```

```{r fig.width=4, fig.height=3, message=FALSE}
sigGene <- "Sell"
plotSmoothers(sce, counts = counts(sce), gene = sigGene) + ggtitle(sigGene) + scale_color_manual(values =c("#33A02C", "#B2DF8A"))
ggsave(filename = paste("../Figures/TradeSeq_plot_IM_lineages_", sigGene, ".pdf"), 
       width = 4, height = 3)
```
```{r fig.width=4, fig.height=3, message=FALSE}
sigGene <- "Ccr2"
plotSmoothers(sce, counts = counts(sce), gene = sigGene) + ggtitle(sigGene) + scale_color_manual(values =c("#33A02C", "#B2DF8A"))
ggsave(filename = paste("../Figures/TradeSeq_plot_IM_lineages_", sigGene, ".pdf"), 
       width = 4, height = 3)
```

```{r fig.width=4, fig.height=3, message=FALSE}
sigGene <- "Itga1"
plotSmoothers(sce, counts = counts(sce), gene = sigGene) + ggtitle(sigGene) + scale_color_manual(values =c("#33A02C", "#B2DF8A"))
ggsave(filename = paste("../Figures/TradeSeq_plot_IM_lineages_", sigGene, ".pdf"), 
       width = 4, height = 3)
```

```{r fig.width=4, fig.height=3, message=FALSE}
sigGene <- "Trem1"
plotSmoothers(sce, counts = counts(sce), gene = sigGene) + ggtitle(sigGene) + scale_color_manual(values =c("#33A02C", "#B2DF8A"))
ggsave(filename = paste("../Figures/TradeSeq_plot_IM_lineages_", sigGene, ".pdf"), 
       width = 4, height = 3)
```
```{r fig.width=4, fig.height=3, message=FALSE}
sigGene <- "Cebpb"
plotSmoothers(sce, counts = counts(sce), gene = sigGene) + ggtitle(sigGene) + scale_color_manual(values =c("#33A02C", "#B2DF8A"))
ggsave(filename = paste("../Figures/TradeSeq_plot_IM_lineages_", sigGene, ".pdf"), 
       width = 4, height = 3)
```

```{r fig.width=4, fig.height=3, message=FALSE}
sigGene <- "Lgals3"
plotSmoothers(sce, counts = counts(sce), gene = sigGene) + ggtitle(sigGene) + scale_color_manual(values =c("#33A02C", "#B2DF8A"))
ggsave(filename = paste("../Figures/TradeSeq_plot_IM_lineages_", sigGene, ".pdf"), 
       width = 4, height = 3)
```

```{r fig.width=4, fig.height=3, message=FALSE}
sigGene <- "Cd81"
plotSmoothers(sce, counts = counts(sce), gene = sigGene) + ggtitle(sigGene) + scale_color_manual(values =c("#33A02C", "#B2DF8A"))
ggsave(filename = paste("../Figures/TradeSeq_plot_IM_lineages_", sigGene, ".pdf"), 
       width = 4, height = 3)
```

```{r fig.width=4, fig.height=3, message=FALSE}
sigGene <- "Tnfaip3"
plotSmoothers(sce, counts = counts(sce), gene = sigGene) + ggtitle(sigGene) + scale_color_manual(values =c("#33A02C", "#B2DF8A"))
ggsave(filename = paste("../Figures/TradeSeq_plot_IM_lineages_", sigGene, ".pdf"), width = 4, height = 3)
```
```{r fig.width=4, fig.height=3, message=FALSE}
sigGene <- "Nfkbiz"
plotSmoothers(sce, counts = counts(sce), gene = sigGene) + ggtitle(sigGene) + scale_color_manual(values =c("#33A02C", "#B2DF8A"))
ggsave(filename = paste("../Figures/TradeSeq_plot_IM_lineages_", sigGene, ".pdf"), width = 4, height = 3)
```

```{r fig.width=4, fig.height=3, message=FALSE}
sigGene <- "Fgfr1"
plotSmoothers(sce, counts = counts(sce), gene = sigGene) + ggtitle(sigGene) + scale_color_manual(values =c("#33A02C", "#B2DF8A"))
ggsave(filename = paste("../Figures/TradeSeq_plot_IM_lineages_", sigGene, ".pdf"), width = 4, height = 3)
```
```{r fig.width=4, fig.height=3, message=FALSE}
sigGene <- "Aif1"
plotSmoothers(sce, counts = counts(sce), gene = sigGene) + ggtitle(sigGene) + scale_color_manual(values =c("#33A02C", "#B2DF8A"))
ggsave(filename = paste("../Figures/TradeSeq_plot_IM_lineages_", sigGene, ".pdf"), width = 4, height = 3)
```

```{r fig.width=4, fig.height=3, message=FALSE}
sigGene <- "Jun"
plotSmoothers(sce, counts = counts(sce), gene = sigGene) + ggtitle(sigGene) + scale_color_manual(values =c("#33A02C", "#B2DF8A"))
ggsave(filename = paste("../Figures/TradeSeq_plot_IM_lineages_", sigGene, ".pdf"), width = 4, height = 3)
```
```{r fig.width=4, fig.height=3, message=FALSE}
sigGene <- "Cd83"
plotSmoothers(sce, counts = counts(sce), gene = sigGene) + ggtitle(sigGene) + scale_color_manual(values =c("#33A02C", "#B2DF8A"))
ggsave(filename = paste("../Figures/TradeSeq_plot_IM_lineages_", sigGene, ".pdf"), width = 4, height = 3)
```
```{r fig.width=4, fig.height=3, message=FALSE}
sigGene <- "H2-Eb1"
plotSmoothers(sce, counts = counts(sce), gene = sigGene) + ggtitle(sigGene) + scale_color_manual(values =c("#33A02C", "#B2DF8A"))
ggsave(filename = paste("../Figures/TradeSeq_plot_IM_lineages_", sigGene, ".pdf"), width = 4, height = 3)
```

```{r fig.width=4, fig.height=3, message=FALSE}
sigGene <- "C1qc"
plotSmoothers(sce, counts = counts(sce), gene = sigGene) + ggtitle(sigGene) + scale_color_manual(values =c("#33A02C", "#B2DF8A"))
ggsave(filename = paste("../Figures/TradeSeq_plot_IM_lineages_", sigGene, ".pdf"), width = 4, height = 3)
```

```{r fig.width=4, fig.height=3, message=FALSE}
sigGene <- "Folr2"
plotSmoothers(sce, counts = counts(sce), gene = sigGene) + ggtitle(sigGene) + scale_color_manual(values =c("#33A02C", "#B2DF8A"))
ggsave(filename = paste("../Figures/TradeSeq_plot_IM_lineages_", sigGene, ".pdf"), width = 4, height = 3)
```

```{r fig.width=4, fig.height=3, message=FALSE}
sigGene <- "Lyve1"
plotSmoothers(sce, counts = counts(sce), gene = sigGene) + ggtitle(sigGene) + scale_color_manual(values =c("#33A02C", "#B2DF8A"))
ggsave(filename = paste("../Figures/TradeSeq_plot_IM_lineages_", sigGene, ".pdf"), width = 4, height = 3)
```
```{r fig.width=4, fig.height=3, message=FALSE}
sigGene <- "Cd38"
plotSmoothers(sce, counts = counts(sce), gene = sigGene) + ggtitle(sigGene) + scale_color_manual(values =c("#33A02C", "#B2DF8A"))
ggsave(filename = paste("../Figures/TradeSeq_plot_IM_lineages_", sigGene, ".pdf"), width = 4, height = 3)
```

```{r fig.width=4, fig.height=3, message=FALSE}
sigGene <- "Mmp9"
plotSmoothers(sce, counts = counts(sce), gene = sigGene) + ggtitle(sigGene) + scale_color_manual(values =c("#33A02C", "#B2DF8A"))
ggsave(filename = paste("../Figures/TradeSeq_plot_IM_lineages_", sigGene, ".pdf"), width = 4, height = 3)
```

```{r fig.width=4, fig.height=3, message=FALSE}
sigGene <- "Cd63"
plotSmoothers(sce, counts = counts(sce), gene = sigGene) + ggtitle(sigGene) + scale_color_manual(values =c("#33A02C", "#B2DF8A"))
ggsave(filename = paste("../Figures/TradeSeq_plot_IM_lineages_", sigGene, ".pdf"), width = 4, height = 3)
```



```{r fig.width=4, fig.height=3, message=FALSE}
sigGene <- "H2-DMb2"
plotSmoothers(sce, counts = counts(sce), gene = sigGene) + ggtitle(sigGene) + scale_color_manual(values =c("#33A02C", "#B2DF8A"))
ggsave(filename = paste("../Figures/TradeSeq_plot_IM_lineages_", sigGene, ".pdf"), width = 4, height = 3)
```

```{r fig.width=4, fig.height=3, message=FALSE}
sigGene <- "H2-Oa"
plotSmoothers(sce, counts = counts(sce), gene = sigGene) + ggtitle(sigGene) + scale_color_manual(values =c("#33A02C", "#B2DF8A"))
ggsave(filename = paste("../Figures/TradeSeq_plot_IM_lineages_", sigGene, ".pdf"), width = 4, height = 3)
```

```{r fig.width=4, fig.height=3, message=FALSE}
sigGene <- "Wnt11"
plotSmoothers(sce, counts = counts(sce), gene = sigGene) + ggtitle(sigGene) + scale_color_manual(values =c("#33A02C", "#B2DF8A"))
ggsave(filename = paste("../Figures/TradeSeq_plot_IM_lineages_", sigGene, ".pdf"), width = 4, height = 3)
```

```{r fig.width=4, fig.height=3, message=FALSE}
sigGene <- "Il1b"
plotSmoothers(sce, counts = counts(sce), gene = sigGene) + ggtitle(sigGene) + scale_color_manual(values =c("#33A02C", "#B2DF8A"))
ggsave(filename = paste("../Figures/TradeSeq_plot_IM_lineages_", sigGene, ".pdf"), width = 4, height = 3)
```

```{r fig.width=4, fig.height=3, message=FALSE}
sigGene <- "Dpp4"
plotSmoothers(sce, counts = counts(sce), gene = sigGene) + ggtitle(sigGene) + scale_color_manual(values =c("#33A02C", "#B2DF8A"))
ggsave(filename = paste("../Figures/TradeSeq_plot_IM_lineages_", sigGene, ".pdf"), width = 4, height = 3)
```

# Session information 

R sesssion: 

```{r}
sessionInfo()
```


# References



