---
subtitle: "2-scRNAseq mapping and counts"
# bibliography: bib.bib
link-citations: yes
date: "`r format (Sys.time(), format='%Y-%m-%d %H:%M:%S %z')`"
output: 
  pdf_document:
    pandoc_args: 
      - '../common.yaml'
      - --listings
    includes: 
      in_header: '../preamble.tex'
    toc: true
    number_sections: true
    fig_width: 7
    fig_height: 6
    fig_caption: true
    df_print: tibble
---

```{r include=FALSE}
knitr::opts_chunk$set(message = FALSE, warning = FALSE)
```

\newpage

# Description

The Cell Ranger (v3.0.2) application (10x Genomics) was then used to demultiplex the BCL files into FASTQ files (cellranger mkfastq), to perform alignment (to Cell Ranger mouse genome references GRCm38/Ensembl97), barcode filtering, UMI counting and to produce gene—barcode matrices (cellranger count).

# CellRanger count from fastq

The script below was used to do the mapping and counting with Cellranger. 

```{bash eval=FALSE}
#!/usr/bin/env bash

	cellranger count	--id=${id} \
				--fastqs="${FastqDir}/${id}" \
				--transcriptome=$REF

```

Here, `$id`, `$FastqDir` and `$REF` are the sample ID, the directory containing fastq files and transcriptome reference directory (mouse genome references GRCm38/Ensembl97). 

# Session information

```{r}
sessionInfo()
```

