---
subtitle: "6-Merge two experiments and plot presentation"
# bibliography: bib.bib
link-citations: yes
date: "`r format (Sys.time(), format='%Y-%m-%d %H:%M:%S %z')`"
output: 
  pdf_document:
    pandoc_args: 
      - '../common.yaml'
      - --listings
    includes: 
      in_header: '../preamble.tex'
    toc: true
    number_sections: true
    fig_width: 7
    fig_height: 6
    fig_caption: true
    df_print: tibble
---

```{r include=FALSE}
knitr::opts_chunk$set(message = FALSE, warning = FALSE)
```

\newpage

# Description

To obtain a dynamic expression profile from refilled monocytes to IM differentiation, we analyzed lung monocytes and IM of IM-DTR mice at different time points of diphtheria toxin (DT) treatment. 

Here, we combined the monocytes/IM data from two experiements (raw data and processed data could be found under accession number GSE193894 and GSE193896 in NCBI GEO). We illustrated the different cell types in a 3-dimension plot and found a transit population bridging classical monocytes and IM. 

# Load data and packages

```{r message=FALSE, warning=FALSE}
# packages
library(Seurat)
library(ggplot2)
```

```{r eval=FALSE}
# data
imdtr1 <- readRDS("../3-scRNAseq_initiation_the_IMs_in_Day4_depletion/results.withSingleR.Rds")
imdtr2 <- readRDS("../5-scRNAseq_initiation_the_IMs_refilling_during_post_deletion/immune.cellType_filtered.withSingleR.seuratObject.Rds")
```

```{r include=FALSE, message=FALSE, warning=FALSE}
# data
imdtr1 <- readRDS("../../../IM_DTR_exp1/analyses/20210114_cell_id/results.withSingleR.Rds")
imdtr2 <- readRDS("../../../IM_DTR_exp2/analyses/20210409_immune_cell_id/immune.cellType_filtered.withSingleR.seuratObject.Rds")
```

# Merge Exp1 and Exp2

## Prepare metadata: 
```{r}
imdtr1$orig.ident <- "IM-DTR1"
imdtr2$orig.ident <- "IM-DTR2"

imdtr1$cell.type0 <- "CD45+"

imdtr1$cell.type2 <- imdtr1$RNA_snn_res.0.3

imdtr1$treatment <- ifelse(imdtr1$treatment=="Cre+", "4d", no = "0d")
#imdtr2$treatment <- sub(imdtr2$treatment, pattern = "HT[0-9]-", replacement = "")
old <- unique(imdtr2$treatment)
new <- c("2d", "0.5d", "1d")
imdtr2$treatment[imdtr2$treatment %in% old] <- new[match(imdtr2$treatment, old)]
```

Now merge: 
```{r}
imdtr3 <- merge(imdtr1, imdtr2, project = "IM-DTR")
```

## Normalize and process: 
```{r message=FALSE, warning=FALSE}
imdtr3 <- NormalizeData(imdtr3)
imdtr3 <- FindVariableFeatures(imdtr3, selection.method = "vst", nfeatures = 2000)
imdtr3 <- ScaleData(imdtr3, features = rownames(imdtr3))
```

```{r message=FALSE, warning=FALSE}
imdtr3 <- RunPCA(imdtr3, features = VariableFeatures(imdtr3))
```

```{r message=FALSE, warning=FALSE}
imdtr3 <- JackStraw(imdtr3, num.replicate = 100)
imdtr3 <- ScoreJackStraw(imdtr3, dims = 1:20)
JackStrawPlot(imdtr3, dims = 1:20)
```
*Samples from Exp1 and Exp2 are so similar. * 

Remove old calculated neighbor (snn): 
```{r}
meta.names <- names(imdtr3@meta.data)
old.snn <- meta.names[startsWith(meta.names, "RNA_snn") | startsWith(meta.names, "res") | startsWith(meta.names, "Clusternames") | startsWith(meta.names, "ClusterNames")]
for (i in old.snn) { 
  imdtr3[[i]] <- NULL
}
```

```{r message=FALSE, warning=FALSE}
imdtr3 <- RunPCA(imdtr3, features = VariableFeatures(imdtr3))
```

```{r}
PCAPlot(imdtr3, group.by = "treatment")
```

```{r}
PCAPlot(imdtr3, group.by = "orig.ident")
```
*Samples from Exp1 and Exp2 are so similar. * 

## Non-linear reduction:

```{r}
results <- RunTSNE(imdtr3, dims = 1:6)
```

Calculate UMAP in 3D:

```{r}
imdtr3.for3d <- RunUMAP(imdtr3, dims = 1:6, n.components = 3L)
```

## Presentation in 3D

```{r}
library(rgl)
library(RColorBrewer)

umap_1 <- imdtr3.for3d@reductions$umap@cell.embeddings[,1]
umap_2 <- imdtr3.for3d@reductions$umap@cell.embeddings[,2]
umap_3 <- imdtr3.for3d@reductions$umap@cell.embeddings[,3]

plot3d(x = umap_1, y = umap_2, z = umap_3, col = factor(imdtr3.for3d$cell.type2, labels = brewer.pal(length(unique(imdtr3.for3d$cell.type2)), name = "Paired")), size = 3)
rglwidget()
```
*3d plot shows only one bridge exists. *

## Cluster cells

```{r message=FALSE, warning=FALSE}
imdtr3.for3d <- FindNeighbors(imdtr3.for3d, dims = 1:10)
imdtr3.for3d <- FindClusters(imdtr3.for3d, resolution = 0.23) 
```

```{r fig.height=4, fig.width=12}
Idents(imdtr3.for3d) <- "RNA_snn_res.0.23"
pal <- brewer.pal(length(levels(imdtr3.for3d)), name = "Paired")
DimPlot(imdtr3.for3d, label = TRUE, split.by = "treatment", cols = pal, reduction = "umap")
```
*The cluster 6 could be DCs*

```{r fig.width=5, fig.height=4}
source("../R/SeuratFreqTable.R")
freq.celltype.list <- list(
  `noTreatment` = Seurat2CellFreqTable(subset(imdtr3.for3d, subset = treatment == "0d"), slotName = "RNA_snn_res.0.23"),   
  `12hDT` = Seurat2CellFreqTable(subset(imdtr3.for3d, subset = treatment == "0.5d"), slotName = "RNA_snn_res.0.23"),  
  `24hDT` = Seurat2CellFreqTable(subset(imdtr3.for3d, subset = treatment == "1d"), slotName = "RNA_snn_res.0.23"),
  `48hDT` = Seurat2CellFreqTable(subset(imdtr3.for3d, subset = treatment == "2d"), slotName = "RNA_snn_res.0.23"),
  `96hDT` = Seurat2CellFreqTable(subset(imdtr3.for3d, subset = treatment == "4d"), slotName = "RNA_snn_res.0.23")
)

source("../R/barChart.R")
barChart(freq.celltype.list) + labs(fill = "Cluster") + scale_fill_manual(values = pal)
```

# Population characterization

```{r fig.width=8, fig.height=15}
library(dplyr)
all_cluster.markers <- FindAllMarkers(imdtr3.for3d, verbose = FALSE)
top20 <- all_cluster.markers %>% group_by(cluster) %>% top_n(n = 20, wt = avg_log2FC)
DoHeatmap(imdtr3.for3d, features = top20$gene, group.colors = pal) + NoLegend()
```
Cluster 0: ClaMo
Cluster 1: PatMo
Cluster 2: MHCII_IM
Cluster 3: CD206_IM
Cluster 4: Chemoattractant
Cluster 5: Transit
Cluster 6: ? 

## Characterization of Cluster 6

GO enrichment for cluster 6:

```{r}
library(topGO)
library(org.Mm.eg.db)
allSymbols <- unique(AnnotationDbi::select(org.Mm.eg.db, keys(org.Mm.eg.db),columns = "SYMBOL")$SYMBOL)
extended_signatures_cluster6 <- filter(all_cluster.markers, avg_log2FC > 0.5 & -log10(p_val) > 2 & cluster == "6")
myInterestingGenes <- 
  as.character(extended_signatures_cluster6$gene)

geneList <- factor(as.integer(allSymbols %in% myInterestingGenes))
names(geneList) <- allSymbols


sampleGOdata <- new("topGOdata", 
                    description = "Simple session", ontology = "BP",
                    allGenes = geneList, 
                    annot = annFUN.org, mapping = "org.Mm.eg.db", ID = "SYMBOL")
resultFisher <- runTest(sampleGOdata, algorithm = "classic", statistic = "fisher")
GenTable(sampleGOdata, classicFisher = resultFisher, topNodes = 20)
```


```{r}
VlnPlot(imdtr3.for3d, features = "Ccr7")
```

Cluster 0: ClaMo
Cluster 1: PatMo
Cluster 2: MHCII_IM
Cluster 3: CD206_IM
Cluster 4: Dying_IM
Cluster 5: Transit
Cluster 6: DC

change the idents
```{r}
levels(imdtr3.for3d)
```

```{r fig.width=16, fig.height=5}
imdtr3.for3d$cell.type2 <- factor(Idents(imdtr3.for3d), labels = c("Classical_Mono", "Patrolling_Mono", "MHCII_IM", "CD206_IM", "Dying_IM", "Transit", "DC"))


Idents(imdtr3.for3d) <- "cell.type2"
imdtr3.for3d$treatment <- factor(imdtr3.for3d$treatment, levels = c("0d", "0.5d", "1d", "2d", "4d"))
DimPlot(imdtr3.for3d, label = TRUE, split.by = "treatment", cols = pal)
```

```{r fig.width=5, fig.height=4}
freq.celltype.list <- list(
  `0` = Seurat2CellFreqTable(subset(imdtr3.for3d, subset = treatment == "0d"), slotName = "cell.type2"),   
  `0.5` = Seurat2CellFreqTable(subset(imdtr3.for3d, subset = treatment == "0.5d"), slotName = "cell.type2"),  
  `1` = Seurat2CellFreqTable(subset(imdtr3.for3d, subset = treatment == "1d"), slotName = "cell.type2"),
  `2` = Seurat2CellFreqTable(subset(imdtr3.for3d, subset = treatment == "2d"), slotName = "cell.type2"),
  `4` = Seurat2CellFreqTable(subset(imdtr3.for3d, subset = treatment == "4d"), slotName = "cell.type2")
)

barChart(freq.celltype.list) + labs(fill = "Cluster") + scale_fill_manual(values = pal)
```

```{r eval=FALSE}
ggsave(filename = "BarPlot_distribution_celltypes_in_samples.pdf",
       path = "../Figures", 
       height = 4, width = 5, 
       device = "pdf")
```

## Cell-cycle analysis

```{r}
library(cowplot)
data("geneinfo_human", package = "nichenetr")
s.genes <- nichenetr::convert_human_to_mouse_symbols(cc.genes.updated.2019$s.genes)
g2m.genes <- nichenetr::convert_human_to_mouse_symbols(cc.genes.updated.2019$g2m.genes)
imdtr3.for3d <- CellCycleScoring(imdtr3.for3d, s.features = s.genes, g2m.features = g2m.genes, set.ident = FALSE)
DimPlot(imdtr3.for3d, group.by = "Phase")
```

```{r fig.height=5, fig.width=16}
DimPlot(imdtr3.for3d, group.by = "Phase", reduction = "umap", split.by = "treatment")

```

2d plot: 
```{r fig.height=6, fig.width=7}
DimPlot(imdtr3.for3d, cols = pal)
```


```{r eval=FALSE}
saveRDS(imdtr3.for3d, file = "./immune_imdtr3.seuratObject.Rds")
```

# Generate plots

2D plot (component 1 and 2): 

```{r fig.height=3.5, fig.width=5}
DimPlot(imdtr3.for3d, cols = pal, dims = c(1,2))
```

```{r eval=FALSE}
ggsave(filename = "UMAPplot_All_cells_PC1_2.pdf",
       path = "../Figures", 
       height = 3.5, width = 5, 
       device = "pdf")
```

2D plot (component 1 and 3): 

```{r fig.height=3.5, fig.width=5}
DimPlot(imdtr3.for3d, cols = pal, dims = c(1,3))
```

```{r eval=FALSE}
ggsave(filename = "UMAPplot_All_cells_PC1_3.pdf",
       path = "../Figures", 
       height = 3.5, width = 5, 
       device = "pdf")
```

3D plot: 

```{r}
umap_1 <- imdtr3.for3d@reductions$umap@cell.embeddings[,1]
umap_2 <- imdtr3.for3d@reductions$umap@cell.embeddings[,2]
umap_3 <- imdtr3.for3d@reductions$umap@cell.embeddings[,3]
```

```{r}
open3d()
points3d(x = umap_1, y = umap_2, z = umap_3, 
       col = factor(Idents(imdtr3.for3d), labels = brewer.pal(length(unique(imdtr3.for3d$cell.type2)), name = "Paired")), 
       size = 7.5, alpha = 0.8, point_antialias = TRUE,
  line_antialias = TRUE)
#decorate3d(  box = FALSE,axes = FALSE, xlab = NULL, ylab = NULL, zlab = NULL)
#axes3d(edges = c("x--", "y--", "z--"), alpha = 0.5) 
# bbox3d(box=TRUE) 
#title3d(xlab = 'UMAP1', ylab = 'UMAP2', zlab = 'UMAP3', alpha = 0.7)
rgl.bringtotop()
rgl.viewpoint(30, -20, zoom = 0.7)
# rgl.postscript( filename = "../Docs/Figures/test.eps", fmt = "eps", drawText = TRUE )
rglwidget()
#close3d()
```

```{r}
open3d()
points3d(x = umap_1, y = umap_2, z = umap_3, 
       col = factor(Idents(imdtr3.for3d), labels = brewer.pal(length(unique(imdtr3.for3d$cell.type2)), name = "Paired")), 
       size = 7.5, alpha = 0.8, point_antialias = TRUE,
  line_antialias = TRUE)
#decorate3d(  box = FALSE,axes = FALSE, xlab = NULL, ylab = NULL, zlab = NULL)
axes3d(edges = c("x--", "y--", "z--"), alpha = 1) 
# bbox3d(box=TRUE) 
#title3d(xlab = 'UMAP1', ylab = 'UMAP2', zlab = 'UMAP3', alpha = 0.7)
rgl.bringtotop()
rgl.viewpoint(30, -20, zoom = 0.7)
# rgl.postscript( filename = "../Docs/Figures/test.eps", fmt = "eps", drawText = TRUE )
rglwidget()
#close3d()
```


```{r}
open3d()
spheres3d(x = umap_1, y = umap_2, z = umap_3, 
       col = factor(Idents(imdtr3.for3d), labels = brewer.pal(length(unique(imdtr3.for3d$cell.type2)), name = "Paired")), 
       radius = 0.1, alpha = 0.5, 
       # ambient      = "gray",
       #emission     = "gray",
       specular     = "gray",
       shininess = 10)
#decorate3d(  box = FALSE,axes = FALSE, xlab = NULL, ylab = NULL, zlab = NULL)
#axes3d(edges = c("x--", "y--", "z--"), alpha = 0.5) 
# bbox3d(box=TRUE) 
#title3d(xlab = 'UMAP1', ylab = 'UMAP2', zlab = 'UMAP3', alpha = 0.7)
rgl.bringtotop()
rgl.viewpoint(30, -20, zoom = 0.7)
# rgl.postscript( filename = "../Docs/Figures/test.eps", fmt = "eps", drawText = TRUE )
rglwidget()
#close3d()
```

Make snapshots for every sample: 


```{r}
plot3d.byGroup <- function(seuratObj, treat, savePNG=FALSE, ...) {
obj <- subset(seuratObj, subset = treatment == treat)
col.pal <- factor(Idents(obj), labels = brewer.pal(length(unique(obj$cell.type2)), name = "Paired"))


umap_1 <- obj@reductions$umap@cell.embeddings[,1]
umap_2 <- obj@reductions$umap@cell.embeddings[,2]
umap_3 <- obj@reductions$umap@cell.embeddings[,3]

umap1.min <- min(seuratObj@reductions$umap@cell.embeddings[,1])
umap1.max <- max(seuratObj@reductions$umap@cell.embeddings[,1])

umap2.min <- min(seuratObj@reductions$umap@cell.embeddings[,2])
umap2.max <- max(seuratObj@reductions$umap@cell.embeddings[,2])

umap3.min <- min(seuratObj@reductions$umap@cell.embeddings[,3]) 
umap3.max <- max(seuratObj@reductions$umap@cell.embeddings[,3])

# scatterplot3d(x = umap_1, y = umap_2, z = umap_3, # old one
#               color = col.pal, pch = 20, 
#               grid=FALSE, box=TRUE, col.axis="black", 
#               main = treat, 
#               ...)
open3d()
points3d(x = umap_1, y = umap_2, z = umap_3, 
       col = col.pal, 
       size = 7.5, alpha = 0.8, 
       point_antialias = TRUE,
  line_antialias = TRUE)

# decorate3d(xlim = c(umap1.min, umap1.max), 
#            ylim = c(umap2.min, umap2.max), 
#            zlim = c(umap3.min, umap3.max), 
#            xlab = NULL, ylab = NULL, zlab = NULL,
#            box = FALSE,axes = FALSE)
# bbox3d(box=TRUE) 
#title3d(xlab = 'UMAP1', ylab = 'UMAP2', zlab = 'UMAP3', alpha = 0.7)
rgl.bringtotop()
#rgl.viewpoint(3, 180)
# rgl.postscript( filename = "../Docs/Figures/test.eps", fmt = "eps", drawText = TRUE )
rgl.viewpoint(30, -20, zoom = 0.7)

if (savePNG) {
  filename <- file.path("../Docs/Figures", paste(treat, ".png", sep = ""))
  png(filename)
  rglwidget()
  dev.off()
} else (return(rglwidget()))
#close3d()
}
```

```{r fig.height=5, fig.width=5}
treat.list <- c("0.5d","1d","2d","4d","0d")
# everytime, save PNG manually with width = 700, height = 630. 
  plot3d.byGroup(imdtr3.for3d, treat.list[1])
```

```{r}
plot3d.byGroup(imdtr3.for3d, treat.list[2])
```


```{r}
plot3d.byGroup(imdtr3.for3d, treat.list[3])
```

```{r}
plot3d.byGroup(imdtr3.for3d, treat.list[4])
```

```{r}
plot3d.byGroup(imdtr3.for3d, treat.list[5])
```

Heatmap: 
```{r fig.width=5, fig.height=6, warning=FALSE, message=FALSE}
DoHeatmap(imdtr3.for3d, features = top20$gene, group.colors = pal, size = 0) + scale_fill_gradientn(colors = c("dark blue", "white", "red")) +
   theme(axis.text.y = element_text(size = 4))
```

```{r eval=FALSE}
ggsave(filename = "HeatMap_top20_in_cell_types_blue_red.pdf",
       path = "../Figures", 
       height = 6, width = 5, 
       device = "pdf")
```

Dot plot show cell type markers: 

```{r fig.width=9, fig.height=4}
DotPlot(imdtr3.for3d, 
        cols = c("light grey", "red"),
        features = c("Ccr2", "Irf7", "Ly6c2",
                     "Ace", "Nr4a1", "Fcgr4", 
                     "H2-Ab1", "H2-Eb1", "Cd74", 
                     "Lyve1", "Mrc1", "C1qa", 
                     "Tnf", "Bax", "Trp53", 
                     "Mki67", "Top2a", "Cdk1", 
                     "Zbtb46", "Ccr7", "Id2"
                     )) + 
   theme(axis.text.x = element_text(angle = 45, hjust = 1))
```

```{r eval=FALSE}
ggsave(filename = "DotPlot_Markers_in_cell_types.pdf",
       path = "../Figures", 
       height = 4, width = 9, 
       device = "pdf")
```

Cell-cycle score: 
```{r}
# color gradient
col_gradient <- colorRampPalette(c("lightgrey", "red"))(256)

# assign color in scaled cell-cycle score
val <- imdtr3.for3d$G2M.Score
col_ix <- round((val - min(val))/(max(val)-min(val)) * 255, digits = 0)
col_pal <- col_gradient[col_ix + 1 ] # because range will be 0:255
names(col_pal) <- names(col_ix)

# remake order to avoid overlay
umap_1.order_g2m <- umap_1[order(imdtr3.for3d$G2M.Score, decreasing = TRUE)]
umap_2.order_g2m <- umap_2[order(imdtr3.for3d$G2M.Score, decreasing = TRUE)]
umap_3.order_g2m <- umap_3[order(imdtr3.for3d$G2M.Score, decreasing = TRUE)]
col_pal <- col_pal[order(imdtr3.for3d$G2M.Score, decreasing = TRUE)]
```


```{r}
open3d()
points3d(x = umap_1.order_g2m, y = umap_2.order_g2m, z = umap_3.order_g2m, 
       col = col_pal, 
       size = 7.5, alpha = 1, 
       point_antialias = TRUE,
  line_antialias = TRUE)
#decorate3d(  box = FALSE,axes = FALSE, xlab = NULL, ylab = NULL, zlab = NULL)
#axes3d(edges = c("x--", "y--", "z--"), alpha = 0.5) 
# bbox3d(box=TRUE) 
#title3d(xlab = 'UMAP1', ylab = 'UMAP2', zlab = 'UMAP3', alpha = 0.7)
rgl.bringtotop()
rgl.viewpoint(30, -20, zoom = 0.7)
# rgl.postscript( filename = "../Docs/Figures/test.eps", fmt = "eps", drawText = TRUE )
rglwidget()
#close3d()
```

```{r}
open3d()
points3d(x = umap_1, y = umap_2, z = umap_3, 
       col = factor(Idents(imdtr3.for3d), labels = brewer.pal(length(unique(imdtr3.for3d$cell.type2)), name = "Paired")), 
       size = 7.5, alpha = 1, 
       point_antialias = TRUE,
  line_antialias = TRUE)
#decorate3d(  box = FALSE,axes = FALSE, xlab = NULL, ylab = NULL, zlab = NULL)
#axes3d(edges = c("x--", "y--", "z--"), alpha = 0.5) 
# bbox3d(box=TRUE) 
#title3d(xlab = 'UMAP1', ylab = 'UMAP2', zlab = 'UMAP3', alpha = 0.7)
rgl.bringtotop()
rgl.viewpoint(30, -20, zoom = 0.7)
# rgl.postscript( filename = "../Docs/Figures/test.eps", fmt = "eps", drawText = TRUE )
rglwidget()
#close3d()
```

Score: 

```{r fig.width=6, fig.height=4}
VlnPlot(imdtr3.for3d, features = "G2M.Score", cols = pal, pt.size = 0)
```
```{r eval=FALSE}
ggsave(filename = "VlnPlot_G2MScore_in_all_samples.pdf",
       path = "../Figures", 
       height = 4, width = 6, 
       device = "pdf")
```

Csf1 expression: 

```{r fig.width=6, fig.height=4}
VlnPlot(imdtr3.for3d, features = "Csf1", cols = pal, pt.size = 0)
```

```{r eval=FALSE}
ggsave(filename = "VlnPlot_Csf1_in_all_samples.pdf",
       path = "../Figures", 
       height = 4, width = 6, 
       device = "pdf")
```

Csf1r
```{r fig.width=6, fig.height=4}
VlnPlot(imdtr3.for3d, features = "Csf1r", cols = pal, pt.size = 0)
```

```{r eval=FALSE}
ggsave(filename = "VlnPlot_Csf1r_in_all_samples.pdf",
       path = "../Figures", 
       height = 4, width = 6, 
       device = "pdf")
```

# Session information

```{r}
sessionInfo()
```
